### 2022-03-14 _paper.tex
wresting jmlr2e.sty

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections},
  pdfauthor={Nicholas Spyrison; Dianne Cook; Kimbal Marriott},
  pdfkeywords={R, XAI, explanable artificial intelligence, black-box
models, local explanations, grand tour, manual tour, radial tour},
  colorlinks=true,
  linkcolor=red,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Exploring Local Explanations of Nonlinear Models Using Animated
Linear Projections}
\author{Nicholas Spyrison \and Dianne Cook \and Kimbal Marriott}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The increased predictive power of nonlinear models comes at the cost of
interpretability of its terms. This trade-off has led to the emergence
of eXplainable AI (XAI). XAI attempts to shed light on how models use
predictors to arrive at a prediction with \emph{local explanations}, a
point estimate of the linear feature importance in the vicinity of one
instance. These can be considered linear projections and can be further
explored to understand better the interactions between features used to
make predictions across the predictive model surface. Here we describe
interactive linear interpolation used for exploration at any instance
and illustrate with examples with categorical (penguin species,
chocolate types) and quantitative (soccer/football salaries, house
prices) output. The methods are implemented in the \textbf{R} package
\textbf{cheem}, available on CRAN.
\end{abstract}

\hypertarget{sec:intro}{%
\section{Introduction}\label{sec:intro}}

There are different reasons and emphases when considering to fit a
model. Breiman (2001b), reiterated by Shmueli (2010), taxonomize models
based on their purpose. \emph{Explanatory} modeling is done for some
inferential purpose, while \emph{predictive} modeling focuses more
narrowly on the performance of some objective function. The intended use
has important implications for model selection and development. In
explanatory modeling, interpretability is vital for drawing inferential
conclusions. Nonlinear models range from additive models with at least
one polynomial term to more complex machine learning models such as
random forest, support-vector machine, or neural network to name a few
(Breiman 2001a; Boser, Guyon, and Vapnik 1992; Anderson 1995,
receptively).

Nonlinear models have many or complex interactive terms, which cause an
opaqueness to the interpretation of the variables. This difficulty in
interpreting the terms in complex nonlinear models sometimes leads to
them being referred to as black box models. Despite the potentially
better performance of nonlinear models, their use is not without
controversy (O'Neil 2016; Kodiyan 2019). And the loss of interpretation
presents a challenge.

Interpretability is vital for exploring and protecting against potential
biases in any model (e.g.~sex -- Dastin (2018); Duffy (2019), race --
Larson et al. (2016), and age -- Díaz et al. (2018)). For instance,
models regularly pick up on biases in the training data where such
classes correlate with changes in the response variable. This bias is
then built into the model. Variable-level (feature-level)
interpretability of models is essential in evaluating and addressing
such biases.

Another concern is data drift, where a shift in the range of the
explanatory variables (features or predictors). Some nonlinear models
are sensitive to this and do not extrapolate well outside the support of
the training data (Espadoto et al. 2021). Maintaining variable
interpretability is also essential to address issues arising from data
drift.

Explainable Artificial Intelligence (XAI) is an emerging field of
research that tries to increase the interpretability of black-box models
(Adadi and Berrada 2018; Arrieta et al. 2020). A common approach is to
use \emph{local explanations}, which attempt to approximate linear
feature importance at the location of each instance (observation) or the
predictions at a specific point in the data domain. Because these are
point-specific, it is challenging to visualize them to comprehensively
understand a model.

In multivariate data visualization, a \emph{tour} (Asimov 1985; Buja and
Asimov 1986; S. Lee et al. 2021) is a sequence of linear projections of
data onto a lower-dimensional space. Tours are viewed as an animation
over minor changes to the projection basis. Structure in a projection
can then be explored visually to see which features contribute to the
formation of that structure. The intuition is similar to watching the
shadow of a hidden 3D object change as the object is rotated; watching
the shape of the shadow change conveys information about the structure
and features of the object.

There are various types of tours distinguished by the generation of
projection bases. In a \emph{manual} tour (Cook and Buja 1997; Spyrison
and Cook 2020), the path is defined by changing the contribution of a
selected feature. Applying tours to models has been done in a couple of
contexts. Specifically for exploring various statistical model fits and
classification boundaries (Wickham, Cook, and Hofmann 2015), and using
tree- and forest-based approaches as a projection pursuit index to
generate a tour basis paths (Y. D. Lee et al. 2013; da Silva, Cook, and
Lee 2021).

We use the radial manual tour to scrutinize a local explanation in our
proposed approach. Additional interactivity allows the user to identify
an instance of interest then explore its local explanation by changing
feature contribution with the radial tour. The methods are implemented
in R package \textbf{cheem}. Example datasets are provided to illustrate
usage for classification and regression tasks.

Using a radial tour can be compared with counterfactual, what-if
analysis, such as \emph{ceteris paribus} profiles (Biecek 2020). Ceteris
paribus, is Latin for ``other things held constant'' or ``all else
unchanged.'' These profiles show how an instance's prediction would
change from a marginal change in one explanatory feature given that
other features are held constant. It ignores correlations of the
features and imagines a case that was not observed. In contrast, our
approach is a geometric explanation of the factual; it varies
contributions of the features by rotating the basis, a reorientation of
the data object. A constraint in our approach is that the basis must
remain orthonormal. When the contribution of one feature decreases, the
contributions of others necessarily increase such that there is a
complete component in that direction. This also ensures that what is
seen is strictly a low-dimensional projection from high-dimensions and
is thus an interpretable visualization.

The remainder of this paper is organized as follows. The following
Section, @ref(sec:explanations), covers the background of the local
explanation and the traditional visuals produced. Section @ref(sec:tour)
explains the animations of continuous linear projections. Section
@ref(sec:cheemviewer) discusses the visual layout in the graphical user
interface, how it facilitates analysis, data preprocessing, and package
infrastructure. Then Section @ref(sec:casestudies) illustrates the
application to supervised learning with categorical and quantitative
output. Section @ref(sec:cheemdiscussion) concludes with the insights
gained and possible directions to explore in the future.

\hypertarget{sec:explanations}{%
\section{Local explanations}\label{sec:explanations}}

Consider a highly nonlinear model. It can be hard to determine whether
small changes in a feature's value will make a class prediction change
group or identify which features contribute to an extreme residual.
Local explanations shed light on these situations by approximating
linear feature importance in the vicinity of a single instance.

A comprehensive summary of the taxonomy and literature of explanation
techniques is provided in Figure 6 of Arrieta et al. (2020). It includes
a large number of model-specific explanations such as deepLIFT
(Shrikumar et al. 2016; Shrikumar, Greenside, and Kundaje 2017), a
popular recursive method for estimating importance in neural networks.
There are fewer model-agnostic explanations, of which LIME, (Ribeiro,
Singh, and Guestrin 2016) SHAP, (Lundberg and Lee 2017), and their
variants are popular.

These instance-level explanations are used in various ways depending on
the data. In image classification, where pixels correspond to
predictors, saliency maps overlay or offset a heatmap indicating
important pixels (Simonyan, Vedaldi, and Zisserman 2014). For instance,
pixels corresponding to snow may be highlighted when distinguishing if a
picture contains a wolf or husky. In text analysis, word-level
contextual sentiment analysis can be used to highlight the sentiment and
magnitude of influential words (Vanni et al. 2018). In the case of
numeric regression, they are used to explain feature additive
contributions from the model intercept to the instance's prediction
(Ribeiro, Singh, and Guestrin 2016).

SHaply Additive exPlanations (SHAP) quantifies the feature contributions
of one instance by examining the effect of other features on the
predictions. The explanations of SHAP refer to Shapley (1953)'s method
to evaluate an individual's contribution to cooperative games by
assessing this player's performance in the presence or absence of other
players. Strumbelj and Kononenko (2010) introduced SHAP for local
explanations in ML models. The attribution of feature importance depends
on the sequence of the included features. The SHAP values are the mean
contributions over different feature sequences. The approach is related
to partial dependence plots (Molnar 2020), used to explain the effect of
a feature by predicting the response for a range of values on this
feature after fixing the value of all other features to their mean.
Though partial dependence plots are a global approximation of the
feature importance, while SHAP is specific to one instance. It could
also be considered similar to examining the coefficients from all
subsets regression, as described in Wickham, Cook, and Hofmann (2015),
helps to understand the relative importance of each feature in the
context of all other candidate features.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/shap_distr_bd}

}

\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) shows breakdown plots of three sequences of the features. The sequence of the variables impacts the magnitude of their attribution. Panel (b) shows the distribution of attribution for each feature across 25 sequences of predictors, with the mean displayed as a dot for each player. Reaction skills are important for both players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
\end{figure}

Following the use case \emph{Explanatory Model Analysis} (Biecek and
Burzykowski 2021), FIFA data is used to illustrate SHAP. Consider soccer
data from the FIFA 2020 season (Leone 2020). There are 5000 instances of
9 skill measures (after aggregating highly correlated features). A
random forest model is fit regressing player's wages {[}2020 Euros{]}
from their skill measurements. The SHAP values are compared for a star
offensive player (L. Messi) and defensive player (V. van Dijk). The
results are displayed in Figure @ref(fig:shapdistrbd). A difference in
the attribution of the feature importance across the two positions of
the players can be expected. This would be interpreted as how a player's
salary depends on this combination of skill sets. Panel (a) is a
modified breakdown plot (Gosiewska and Biecek 2019) where three
sequences of features are presented, so the two instances can be more
easily compared. The magnitude of the contributions depends on the
sequence in which they appear. Panel (b) shows the differences of the
player's median values of over 25 such sequences. In summary, these
plots highlight how local explanations bring interpretability to a
model, at least in the vicinity of their instances. In this instance,
two players with different positions receive different profiles of
feature importance to explain the prediction of their wages.

For the application, we use \emph{tree SHAP}, a variant of SHAP that
enjoys a lower computational complexity (Lundberg, Erion, and Lee 2018).
Instead of aggregating over sequences of the features, tree SHAP
calculates instance-level feature importance by exploring the structure
of the decision trees. Tree SHAP is only compatible with tree-based
models; random forests are used for illustration. The following section
will use normalized SHAP values as a projection basis (call this the
\emph{attribution projection}) that will be used to explore the
sensitivity of the feature contributions.

\hypertarget{sec:tour}{%
\section{Tours and the radial tour}\label{sec:tour}}

A \emph{tour} enables viewing of high-dimensional data by animating many
linear projections with small incremental changes. It is achieved by
following a path of linear projections (bases) of high-dimensional
space. One of the features of the tour is the object permanence of the
data points; one can track the relative change of instances in time and
gain information about the relationships between points across multiple
features. There are various types of tours that are distinguished by how
the paths are generated (S. Lee et al. 2021; Cook et al. 2008).

The manual tour (Cook and Buja 1997) defines its path by changing a
selected feature's contribution to a basis to allow the feature to
contribute more or less to the projection. The requirement constrains
the contribution of all other features that a basis needs to be
orthonormal (column correspond to vectors, with unit length, and
orthogonal to each other). The manual tour is primarily used to assess
the importance of a feature to structure visible in a projection. It
also lends itself to pre-computation queued in advance or computed
on-the-fly for human-in-the-loop analysis (Karwowski 2006).

A version of the manual tour called a \emph{radial tour} is implemented
in Spyrison and Cook (2020) and forms the basis of the new work. In a
radial tour, the selected feature can change its magnitude of
contribution but not its angle; it must move along the direction of its
original contribution. The implementation allows for pre-computation and
interactive re-calculation to focus on a different feature.

\hypertarget{sec:cheemviewer}{%
\section{The cheem viewer}\label{sec:cheemviewer}}

To explore the local explanations, coordinated views (Roberts 2007)
(also known as ensemble graphics, Unwin and Valero-Mora 2018) are
provided in the \emph{cheem viewer} application. There are two primary
plots: the \textbf{global view} to give the context of all of the SHAP
values and the \textbf{radial tour view} to explore the local
explanations with user-controlled rotation. There are numerous user
inputs, including feature selection for the radial tour and instance
selection for making comparisons. There are different plots used for the
categorical and quantitative responses. Figures
@ref(fig:classificationcase) and @ref(fig:regressioncase) are
screenshots showing the cheem viewer for the two primary tasks:
classification (categorical response) and regression (quantitative
response).

\hypertarget{global-view}{%
\subsection{Global view}\label{global-view}}

The global view provides a context of all instances and facilitates the
exploration of the separability of the data- and attribution-spaces.
Both of these spaces are of dimension \(n\times p\), where \(n\) is the
number of instances and \(p\) is the number of predictors. The
attribution space corresponds to the local explanations for each
instance, each a vector of \(p\) values.

Visualization is provided by the first two principal components of the
data (left) and the attribution (middle) spaces. These single 2D
projections will not reveal all of the structure of higher-dimensional
space, but they are helpful visual summaries. In addition, a plot of the
observed against predicted response values is also provided (Figures
@ref(fig:classificationcase)b, @ref(fig:regressioncase)a) to help
identify instances poorly predicted by the model. For classification
tasks, misclassified instances are circled in red. Linked brushing
between the plots is provided, and a tabular display of selected points
helps to facilitate exploration of the spaces and the model (shown in
Figures @ref(fig:regressioncase)d).

While the comparison of these spaces is interesting, the primary purpose
of the global view is to enable the selection of instances to explore
the local explanations. The projection attribution of the primary
instance (PI) is examined and typically viewed with an optional
comparison instance (CI). These instances are highlighted as asterisk
and \(\times\), respectively.

\hypertarget{radial-tour}{%
\subsection{Radial tour}\label{radial-tour}}

The local explanations for all observations are normalized (squared sum
of values adds to 1), and thus, the relative importance of features can
be compared across all instances. These are depicted as vertical
parallel coordinate plots (Ocagne 1885) on the basis biplot (Gabriel
1971). 1D biplot display depicts the values of the current basis as
bars. The parallel coordinate overlays lines connecting one instance's
feature attribution (Figures @ref(fig:classificationcase)e and
@ref(fig:regressioncase)e). The attribution projections of the PI and CI
are shown as dashed and dotted lines. From this plot, the range and
density of the importance across all instances can be interpreted. For
classification, one would look at differences between groups on any
feature. For example, Figure @ref(fig:classificationcase)e suggests that
\texttt{bl} is important for distinguishing the green class from the
other two. For regression, one might generally observe which features
have low values for all instances (not important), for example,
\texttt{BMI} and \texttt{pwr} in Figure @ref(fig:regressioncase)e, and
which have a range of high and low values (e.g.~\texttt{off},
\texttt{def}) suggesting they are important for some instances and not
important for other instances.

The overlaid bars on the parallel coordinate plot represent the
attribution projection of the PI. (Remember that the PI is interactively
selected from the global view). The attribution projection is
approximates the feature importance for predicting this instance. This
combination of features best explains the difference between the mean
response and an instance's predicted value. It is not an indication of
the local shape of the model surface. That is, it is not some indication
of the tangent to the curve at this point.

The attribution projection of the PI is the initial 1D basis in a radial
tour, displayed as a density plot for a categorical response (Figure
@ref(fig:classificationcase)f) and as scatterplots for a quantitative
response (Figure @ref(fig:regressioncase)f). The PI and CI are indicated
by vertical dashed and dotted lines, respectively. The radial tour
varies the contribution of the selected feature between 0-1. This is
viewed as an animation of the projections from many intermediate bases.
Doing so tests the sensitivity of structure (class separation or
strength of relationship) to the feature's contribution. For
classification, if the separation between classes diminishes when the
feature contribution is reduced, this suggests that the feature is
important for class separation. For regression, if the relationship
scatterplot weakens when the feature contribution is reduced, indicating
that the feature is important for accurately predicting the response.

\hypertarget{classification-task}{%
\subsection{Classification task}\label{classification-task}}

Selecting a misclassified instance as PI and a correctly classified
point nearby in data space as CI makes it easier to examine the features
most responsible for the error. The global view (Figure
@ref(fig:classificationcase)c) displays the model confusion matrix. The
radial tour is 1D and displays as density where color indicates class.
An animation slider enables the user to vary the contribution of each
feature to explore the sensitivity of the separation to that feature.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}

}

\caption{Overview of the cheem viewer for classification tasks. Global view inputs, (a), set the PI, CI, and color statistic. Global view, (b) PC1 by PC2 approximations of the data space and attribution space. (c) prediction by observed y (visual of the confusion matrix for classification tasks). Points are colored by predicted class, and red circles indicate misclassified instances. Radial tour inputs (d) select features to include and which feature is changed in the tour. (e) shows a parallel coordinate display of the distribution of the feature attributions while bars depict contribution for the current basis. The black bar is the variable being changed in the radial tour. Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
\end{figure}

\hypertarget{regression-task}{%
\subsection{Regression task}\label{regression-task}}

Selecting an inaccurately predicted instance as PI and an accurately
predicted instance with similar feature values as CI is a helpful way to
understand how the model is failing or not. The global view (Figure
@ref(fig:regressioncase)a) shows a scatterplot of the observed vs
predicted values, which should exhibit a strong relationship if the
model is a good fit. The points can be colored by a statistic, residual,
a measure of outlyingness (log Mahalanobis distance), or correlation aid
in understanding the structure identified in these spaces.

In the radial tour view, the observed response and the residuals
(vertical) are plotted against the attribution projection of the PI
(horizontal). The attribution projection can be interpreted similarly to
the predicted value from the global view plot. It represents a linear
combination of the features, and a good fit would be indicated when
there is a strong relationship with the observed values. This can be
viewed as a local linear approximation if the fitted model is nonlinear.
As the contribution of a feature is varied, if the value of the PI does
not change much, it would indicate that the prediction for this instance
is NOT sensitive to that feature. Conversely, if the predicted value
varies substantially, the prediction is very sensitive to that feature,
suggesting that the feature is very important for the PI's prediction.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}

}

\caption{Overview of the cheem viewer for regression  and illustration of interactive features. Panel (a) PCA of the data and attributions spaces, (b) residual plot, predictions by observed values. Four selected points are highlighted in the PC spaces and tabularly displayed. Coloring on a statistic (c) highlights structure organized in the attribution space. Interactive tabular display (d) populates when instances are selected. Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the feature attribution from all observations, and horizontal bars show the contribution to the current basis. Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed y and residuals (middle and right).}\label{fig:regressioncase}
\end{figure}

\hypertarget{interactive-features}{%
\subsection{Interactive features}\label{interactive-features}}

The application has several reactive inputs that affect the data used,
aesthetic display, and tour manipulation. These reactive inputs make the
software flexible and extensible (Figure @ref(fig:classificationcase)a
\& d). The application also has more exploratory interactions to help
link points across displays, reveal structures found in different
spaces, and access the original data.

A tooltip displays instance number/name and classification information
while the cursor hovers over a point. Linked brushing allows the
selection of points (left click and drag) where those points will be
highlighted across plots (Figure @ref(fig:classificationcase)a \& b).
The information corresponding to the selected points is populated on a
dynamic table (Figure @ref(fig:classificationcase)d). These interactions
aid exploration of the spaces and, finally, the identification of
primary and comparison instances.

\hypertarget{preprocessing}{%
\subsection{Preprocessing}\label{preprocessing}}

It is vital to mitigate the render time of visuals, especially when
users may want to iterate many explorations. All computational
operations should be prepared before runtime. The work remaining when an
application is run solely reacts to inputs and rendering visuals and
tables. Below discusses the steps and details of the reprocessing.

\begin{itemize}
\tightlist
\item
  \textbf{Data:} predictors and response are unscaled complete numerical
  matrix. Most models and local explanations are scale-invariant.
\item
  \textbf{Model:} any model and compatible explanation could be explored
  with this method. Currently, random forest models are applied via the
  package \textbf{randomForest} (Liaw and Wiener 2002), compatibility
  tree SHAP. Modest hyperparameters are used, namely: 125 trees, number
  variables randomly sampled at each split, mtry = \(\sqrt{p}\) or
  \(p/3\) for classification and regression, and minimum size of
  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
  classification and regression.
\item
  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
  observation using the package \textbf{treeshap} (Kominsarczyk et al.
  2021). This implementation aggregates exhaustively over all trees'
  attribution and does not fit interactions of variables.
\item
  \textbf{Cheem viewer:} after the model and full explanation space are
  calculated, each variable is scaled by standard deviations away from
  the mean to achieve common support for visuals. Statistics for mapping
  to color are computed on the scaled spaces.
\end{itemize}

The time to preprocess the data will vary significantly with the
complexity of the model and local explanation. For reference, the FIFA
data contained 5000 instances of nine explanatory features took 2.5
seconds to fit a random forest model of modest hyperparameters.
Extracting the tree SHAP values of each instance took 270 seconds total.
PCA and statistics of the features and attributions took 2.8 seconds.
These runtimes were from a non-parallelized R session on a modern
laptop, but suffice to say that most of the time will be spent on the
local attribution. An increase in model complexity or data
dimensionality will quickly become an obstacle. Its reduced
computational complexity makes tree SHAP an excellent candidate to
start. (Alternatively, the package \textbf{fastshap} (Greenwell 2020)
claims extremely low runtimes, attributed to fewer calls to the
prediction function, partial implementation in C++, and efficient use of
logical subsetting.)

\hypertarget{sec:infrastructure}{%
\subsection{Package infrastructure}\label{sec:infrastructure}}

The above-described method and application are implemented as an
open-source \textbf{R} package, \textbf{cheem} available on
\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Preprocessing was
facilitated with models created via \textbf{randomForest} (Liaw and
Wiener 2002) and explanations calculated with \textbf{treeshap}
(Kominsarczyk et al. 2021). The application was made with \textbf{shiny}
(Chang et al. 2021). The tour visual is built with \textbf{spinifex}
(Spyrison and Cook 2020). Both views are created first with
\textbf{ggplot2} (Wickham 2016) and then rendered as interactive
\texttt{html} widgets with \textbf{plotly} (Sievert 2020).
\textbf{DALEX} (Biecek 2018) and the free ebook, \emph{Explanatory Model
Analysis} (Biecek and Burzykowski 2021) were a boon to understanding
local explanations and how to apply them.

\hypertarget{installation-and-getting-started}{%
\subsection{Installation and getting
started}\label{installation-and-getting-started}}

The package can be installed from CRAN, and the shiny app can be run,
using the following \textbf{R} code:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{install.packages}\NormalTok{(}\StringTok{"cheem"}\NormalTok{, }\AttributeTok{dependencies =} \ConstantTok{TRUE}\NormalTok{)}
\FunctionTok{library}\NormalTok{(}\StringTok{"cheem"}\NormalTok{)}
\FunctionTok{run\_app}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

Alternatively,

\begin{itemize}
\tightlist
\item
  a version of the cheem viewer shiny app can be directly accessed at
  \href{https://ebsmonash.shinyapps.io/cheem_initial/}{ebsmonash.shinyapps.io/cheem\_initial/}.
\item
  the development version of the package is available at
  \url{https://github.com/nspyrison/cheem}, and
\item
  documentation of the package can be found at
  \url{https://nspyrison.github.io/cheem/}.
\end{itemize}

Follow the examples provided with the package to compute the local
explainers (using \texttt{?cheem\_ls}). The application expects the
output returned by \texttt{cheem\_ls()}, saved to an .rds file with
\texttt{saveRDS()}, to be uploaded.

\hypertarget{sec:casestudies}{%
\section{Case studies}\label{sec:casestudies}}

To illustrate the cheem method, it is applied to modern datasets, two
classification examples and then two of regression.

\hypertarget{palmer-penguin-species-classification}{%
\subsection{Palmer penguin, species
classification}\label{palmer-penguin-species-classification}}

The Palmer penguins data (Gorman, Williams, and Fraser 2014; Horst,
Hill, and Gorman 2020) was collected on three species of penguins
foraging near Palmer Station, Antarctica. The data was publicly
available to substitute for the overly-used iris data and is quite
similar in form. After removing incomplete instances, there are 333
instances of four physical measurements, bill length (\texttt{bl}), bill
depth (\texttt{bd}), flipper length (\texttt{fl}), body mass
(\texttt{bm}), for this illustration. A random forest model was fit with
species as the response feature.

(ref:casepenguins-cap) Examining the SHAP values for a random forest
model classifying Palmer penguin species. The PI is a Gentoo (purple)
penguin that is misclassified as a Chinstrap (orange), marked as an
asterisk in (a), and the dashed vertical line in (b). The radial view
shows varying the contribution of \texttt{fl} from the initial
attribution projection (b, left), which produces a linear combination
where the PI is more probably (higher density value) a Chinstrap than a
Gentoo (b, right). (The animation of the radial tour is at
\href{https://vimeo.com/666431172}{vimeo.com/666431172}.)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}

}

\caption{(ref:casepenguins-cap)}\label{fig:casepenguins}
\end{figure}

Figure @ref(fig:casepenguins) shows plots from the cheem viewer for
exploring the random forest model on the penguins data. Panel (a) shows
the global view, and panel (b) shows several 1D projections generated
with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
it has been misclassified as a Chinstrap (orange).

(ref:casepenguinsblfl-cap) Checking what is learned from the cheem
viewer. This is a plot of flipper length (\texttt{fl}) and bill length
(\texttt{bl}), where an asterisk highlights the PI. A Gentoo (purple)
misclassified as a Chinstrap (orange). The PI has an unusually small
\texttt{fl} length which is why it is confused with a Chinstrap.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}

}

\caption{(ref:casepenguinsblfl-cap)}\label{fig:casepenguinsblfl}
\end{figure}

There is more separation visible in the attribution space than the data
space, as would be expected. The predicted vs observed plot reveals a
handful of misclassified instances. A Gentoo that has been wrongly
labeled as a Chinstrap is selected for illustration. The PI is a
misclassified point (represented by the asterisk in the global view and
a dashed vertical line in the tour view). The CI is a correctly
classified point (represented by an \(\times\) and a vertical dotted
line).

The radial tour starts from the attribution projection of the
misclassified instance (b, left). The important features identified by
SHAP in the (wrong) prediction for this instance are mostly \texttt{bl}
and \texttt{bd} with small contributions of \texttt{fl} and \texttt{bm}.
This projection is a view where the Gentoo (purple) looks much more
likely for this instance than Chinstrap. That is, this combination of
features is not particularly useful because the PI looks very much like
other Gentoo penguins. To explore this, the radial tour is used to vary
the contribution of flipper length (\texttt{fl}). (In our exploration,
this was the third feature explored. It is typically helpful to explore
the features with more significant contributions, here \texttt{bl} and
\texttt{bd}. Still, when doing this, nothing was revealed about how the
PI differed from other Gentoos). On varying \texttt{fl} as it
contributes increasingly to the projection (b, right), more and more,
this penguin looks like a Chinstrap. This suggests that \texttt{fl}
should be considered an important feature for explaining the (wrong)
prediction.

Figure @ref(fig:casepenguinsblfl) confirms that flipper length
(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
flipper length and body length are plotted, and the PI can be seen to be
closer to the Chinstrap group in these two features, mainly because it
has an unusually low value of flipper length relative to other Gentoos.
From this view, it makes sense that it is a hard instance to account for
as decision trees can only partition only vertical and horizontal lines.

\hypertarget{chocolates-milkdark-chocolate-classification}{%
\subsection{Chocolates, milk/dark chocolate
classification}\label{chocolates-milkdark-chocolate-classification}}

The chocolates dataset consists of 88 instances of ten nutritional
measurements determined from their labels and labeled as either milk or
dark. Dark chocolate is considered healthier than milk. The data were
collected by students during the Iowa State University class STAT503
from nutritional information from the manufacturer's website and
normalized to 100g equivalents. The data is available in the
\textbf{cheem} package. A random forest model is used for the
classification of chocolate types.

It could be interesting to examine the nutritional properties of any
dark chocolates that have been misclassified as milk. A reason to do
this is that a dark chocolate, nutritionally more like milk, should not
be considered a healthy alternative. It is interesting to explore which
nutritional features contribute most to misclassification.

(ref:casechocolates-cap) Examining the local explanation for a PI which
is dark (orange) chocolate incorrectly predicted to be milk (green).
From the attribution projection, this chocolate correctly looks more
like dark than milk, which suggests that the local explanation does not
help understand the prediction for this instance. So, the contribution
of Sugar is varied --- reducing it corresponds primarily with increasing
Fiber. When Sugar is zero, Fiber contributes strongly towards the left.
In this particular view, the PI is closer to the bulk of the milk
chocolates, suggesting that the prediction put a lot of importance on
Fiber. This chocolate is a rare dark chocolate without any Fiber leading
to it being mistaken for a milk chocolate. (A video of the tour
animation can be found at
\href{https://vimeo.com/666431143}{vimeo.com/666431143}.)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates}

}

\caption{(ref:casechocolates-cap)}\label{fig:casechocolates}
\end{figure}

This type of exploration is shown in Figure @ref(fig:casechocolates),
where a chocolate labeled dark but predicted to be milk is chosen as the
PI (instance 22). It is compared with a CI that is a correctly
classified dark chocolate (instance 7). The PCA plot and the tree SHAP
PCA plots (a) show a big difference between the two chocolate types but
with confusion for a handful of instances. The misclassifications are
more apparent in the observed vs predicted plot and can be seen to be
mistaken in both ways: milk to dark and dark to milk.

The attribution projection for chocolate 22 suggests that Fiber, Sugars,
and Calories are most responsible for its incorrect prediction. The way
to read this plot is to see that Fiber has a large negative value while
Sugars and Calories have reasonably large positive values. In the
density plot, instances on the very left of the display would have high
values of Fiber (matching the negative projection coefficient) and low
values of Sugars and Calories. The opposite would be interpreting a
point with high values in this plot. The dark chocolates (orange) are
primarily on the left, and this is a reason why they are considered to
be healthier: high fiber and low sugar. The density of milk chocolates
is further to the right, indicating that they generally have low fiber
and high sugar.

The PI (dashed line) can be viewed against the CI (dotted line). Now,
one needs to pay attention to the parallel plot of the SHAP values,
which are local to a particular instance, and the density plot, which is
the same projection of all instances as specified by the SHAP values of
the PI. The feature contributions to the two different predictions can
be quickly compared in the parallel coordinate plot. The PI differs with
the comparison primarily on the Fiber feature, which suggests that this
is the reason for the incorrect prediction.

From the density plot, which is the attribution projection corresponding
to the PI, both instances are more like dark chocolates. Varying the
contribution of Sugars and altogether removing it from the projection is
where the difference becomes apparent. When primarily Fiber is examined,
instance 22 looks more like a milk chocolate.

It would also be interesting to explore an inverse misclassification. In
this case, a milk chocolate is selected while it was misclassified as a
dark chocolate. Chocolate 84 is just this case and is compared with a
correctly predicted milk chocolate (instance 71). The corresponding
global view and radial tour frames are shown in Figure
@ref(fig:casechocolatesinverse).

(ref:casechocolatesinverse-cap) Examining the local interpretation for a
PI which is milk (green) chocolate incorrectly predicted to be dark
(orange). In the attribution projection, the PI could be either milk or
dark. Sodium and Fiber have the largest differences in attributed
feature importance, with low values relative to other milk chocolates.
The lack of importance attributed to these variables is suspected of
contributing to the mistake, so the contribution of Sodium is varied. If
Sodium had a larger contribution to the prediction (like in this view).
the PI would look more like other milk chocolates. (A video of the tour
animation can be found at
\href{https://vimeo.com/666431148}{vimeo.com/666431148}.)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates_inverse}

}

\caption{(ref:casechocolatesinverse-cap)}\label{fig:casechocolatesinverse}
\end{figure}

The difference of position in the tree SHAP PCA with the previous case
is quite significant; this gives a higher-level sense that the
attributions should be quite different. Looking at the attribution
projection, this is found to be the case. Previously, Fiber was
essential while it is absent from the attribution in this case.
Conversely, Calories from Fat and Total Fat have high attributions here,
while they were unimportant in the preceding case.

Comparing the attribution with the CI (dotted line), large discrepancies
in Sodium and Fiber are identified. The contribution of Sodium is
selected to be varied. Even in the initial projection, the instance
looks slightly more like its observed milk than predicted dark
chocolate. The misclassification appears least supported when the basis
reaches sodium attribution of typical dark chocolate.

\hypertarget{fifa-wage-regression}{%
\subsection{FIFA, wage regression}\label{fifa-wage-regression}}

The 2020 season FIFA data (Leone 2020; Biecek 2018) contains many skill
measurements of soccer/football players and wage information. Nine
higher-level skill groupings were identified and aggregated from highly
correlated features. A random forest model is fit from these predictors,
regressing player wages {[}2020 euros{]}. The model was fit from 5000
instances before being thinned to 500 players to mitigate occlusion and
render time. Continuing from the exploration in section
@ref(sec:explanations), we are interested to see the difference in
attribution based on the exogenous player position. That is, the model
should be able to use multiple linear profiles to better predict the
wages from different field positions of players despite not having this
information. A leading offensive fielder (L. Messi) is compared with a
top defensive fielder (V. van Dijk). The same instances were used in
figure @ref(fig:shapdistrbd).

(ref:casefifa-cap) Exploring the wages {[}euros{]} relative to skill
measurements in the FIFA 2020 data. Star offensive player (L. Messi) is
the PI and he is compared with a top defensive player (V. van Dijk). The
attribution projection is shown at left, and it can be seen that this
combination of features produces a view where Messi has very high
predicted (and observed) wages. Defense (\texttt{def}) is the chosen
feature to vary. It starts off very low, and as its contribution is
increased (right plot) it can be seen that Messi's predicted wages
decrease dramatically. The increased contribution in defense comes at
the expense of offensive and reaction skills. The interpretation is that
indeed, Messi's high wages are most attributable to his offensive and
reaction skills, as initially provided by the local explanation. (A
video of the animated radial tour can be found at
\href{https://vimeo.com/666431163}{vimeo.com/666431163}.)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_fifa}

}

\caption{(ref:casefifa-cap)}\label{fig:casefifa}
\end{figure}

Figure @ref(fig:casefifa), tests the support of the local explanation.
Offensive and reaction skills (\texttt{off} and \texttt{rct}) are both
crucial to explaining a star offensive player. If either of them were
rotated out, the other would be rotated into the frame, maintaining a
far-right position. However, Increasing the contribution of a variable
with low importance would rotate both features out of the frame.

The contribution from \texttt{def} will be varied to contrast with
offensive skills. As the contribution of defensive skills increases,
Messi's is no longer separated from the group. Players with high values
in defensive skills are now the rightmost points. In terms of what-if
analysis, the difference between the data mean and his predicted wages
would be halved if Messi's tree SHAP attributions at these levels.

\hypertarget{ames-housing-2018-sales-price-regression}{%
\subsection{Ames housing 2018, sales price
regression}\label{ames-housing-2018-sales-price-regression}}

Ames 2018, housing data was subset to North Ames (the neighborhood with
the most house sales). The remaining are 338 house sales. A random
forest model has fitted this price {[}USD{]} with the property features:
Lot Area (\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house
was Built (\texttt{YrB}), Living Area (\texttt{LvA}), number of
Bathrooms (\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total
number of Rooms (\texttt{Rms}), Year the Garage was Built
(\texttt{GYB}), and Garage Area (\texttt{GrA}). Using interaction from
the global view, a house with an extreme negative residual and an
accurate instance with a similar prediction are selected.

(ref:caseames-cap) Exploring an instance with a large residual as the PI
from fitting sales price {[}USD{]} to other variables in the Ames
housing 2018. The sale price of the PI was under-predicted. The local
explanation indicates a sizable attribution to Lot Area (\texttt{LtA}).
The CI has a similar predicted sales price and smaller residual and has
minimal attribution to Lot Area. In the attribution projection, the PI
has a higher sales price than the CI. Reducing the contribution of Lot
Area brings these two prices in line. This suggests if the model did not
consider Lot Area, then the two houses would be quite similar. That is,
the large residual is due to a lack of factoring in the Lot Area for the
prediction of PI's sales price. (A video showing the animation is at
\href{https://vimeo.com/666431134}{vimeo.com/666431134}.)

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_ames2018}

}

\caption{(ref:caseames-cap)}\label{fig:caseames}
\end{figure}

Figure @ref(fig:caseames) selects the house sale 74, a sizable under
prediction with an enormous Lot Area contribution. The CI has a similar
predicted price though the prediction was accurate and gives almost no
attribution to lot size. The attribution projection places instances
with high Living Areas to the right. The contribution of Living Area
contrasts the contribution of this feature. As the contribution of Lot
Area decreases, the predictive power decreases for the PI, while the CI
remains stationary. This large of importance in the Living Area is
relatively uncommon. Boosting tree models may be more resilient to such
an under-prediction as they would up-weighting this residual and force
its inclusion in the final model.

\hypertarget{sec:cheemdiscussion}{%
\section{Discussion}\label{sec:cheemdiscussion}}

There is a clear need to extend the interpretability of black-box
models. This chapter provides a technique that builds on local
interpretations to explore the feature importance local to an instance.
The local interpretations form an attribution projection from which
feature contributions are varied using a radial tour. Several diagnostic
plots are provided to assist with understanding the sensitivity of the
prediction to particular features. A global view shows the data space,
explanation space, residual plot. The user can interactively select
instances to compare, contrast, and study further. Then the radial tour
is used to explore the feature sensitivity identified by the attribution
projection.

This approach has been illustrated using four data examples of random
forest models with the tree SHAP local explanation. In the penguins
example, we showed how the misclassification of a penguin arose due to
it having an unusually small flipper size compared to others of its
species. This was verified by making a follow-up plot of the data. The
chocolates example shows how a dark chocolate was misclassified
primarily due to its attribution to Fiber, and a milk chocolate was
misclassified as dark due to its lowish Sodium value. In the FIFA
example, we show how low Messi's salary would be if it depended on
defensive skills. In the Ames housing data, an inaccurate prediction for
a house was likely due to the Lot Area not being effectively used by the
random forest model.

This analysis is manually intensive and thus only feasible for
investigating a few instances. The approach selects instances that the
model has not done well to predict and compare it with an instance where
it did fit well. The radial tour launches from the attribution
projection to enable exploration of the sensitivity of the prediction to
any feature. It can be helpful to make additional plots of the features
and responses to cross-check interpretations made from the cheem viewer.
This methodology provides an additional tool in the box for studying
model fitting.

An implementation is provided in the open-source \textbf{R} package
\textbf{cheem}, available on
\href{https://CRAN.R-project.org/package=cheem}{CRAN}. Example data sets
are provided, and you can upload your data after model fitting and
computing the local explanations. In theory, this approach would work
with any black-box model, but the implementation currently only
calculates tree SHAP for tree-based models supported by
\textbf{treeshap} (tree based models from \textbf{gbm},
\textbf{lightgbm}, \textbf{randomForest}, \textbf{ranger}, or
\textbf{xgboost} Greenwell et al. 2020; Shi et al. 2022; Liaw and Wiener
2002; Wright and Ziegler 2017; Chen et al. 2021, respectively). Tree
SHAP was selected because of its computational efficiency. The SHAP and
oscillation explanations could be added with the use of the
\texttt{DALEX::explain()} and would be an excellent direction to extend
the work (Biecek 2018; Biecek and Burzykowski 2021).

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We would like to thank Professor Przemyslaw Biecek for his input early
in the project and to the broader \(\text{MI}^\text2\) lab group for the
\textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages.
This research was supported by Australian Government Research Training
Program (RTP) scholarships. Thanks to Jieyang Chong for helping
proofread this article.

The namesake, Cheem, refers to a fictional race of humanoid trees from
Doctor Who lore. \textbf{DALEX} pulls on from that universe, and we
initially apply tree SHAP explanations specific to tree-based models.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-adadi_peeking_2018}{}%
Adadi, Amina, and Mohammed Berrada. 2018. {``Peeking Inside the
Black-Box: A Survey on Explainable Artificial Intelligence ({XAI}).''}
\emph{IEEE Access} 6: 52138--60.

\leavevmode\hypertarget{ref-anderson_introduction_1995}{}%
Anderson, James A. 1995. \emph{An Introduction to Neural Networks}. MIT
press.

\leavevmode\hypertarget{ref-arrieta_explainable_2020}{}%
Arrieta, Alejandro Barredo, Natalia Díaz-Rodríguez, Javier Del Ser,
Adrien Bennetot, Siham Tabik, Alberto Barbado, Salvador García, Sergio
Gil-López, Daniel Molina, and Richard Benjamins. 2020. {``Explainable
{Artificial} {Intelligence} ({XAI}): {Concepts}, Taxonomies,
Opportunities and Challenges Toward Responsible {AI}.''}
\emph{Information Fusion} 58: 82--115.

\leavevmode\hypertarget{ref-asimov_grand_1985}{}%
Asimov, Daniel. 1985. {``The {Grand} {Tour}: A {Tool} for {Viewing}
{Multidimensional} {Data}.''} \emph{SIAM Journal on Scientific and
Statistical Computing} 6 (1): 128--43.
https://doi.org/\url{https://doi.org/10.1137/0906011}.

\leavevmode\hypertarget{ref-biecek_dalex_2018}{}%
Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive
Models in {R}.''} \emph{The Journal of Machine Learning Research} 19
(1): 3245--49.

\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
---------. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus}
{Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.

\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
{Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive}
{Models}}. CRC Press.

\leavevmode\hypertarget{ref-boser_training_1992}{}%
Boser, Bernhard E., Isabelle M. Guyon, and Vladimir N. Vapnik. 1992.
{``A Training Algorithm for Optimal Margin Classifiers.''} In
\emph{Proceedings of the Fifth Annual Workshop on {Computational}
Learning Theory}, 144--52.

\leavevmode\hypertarget{ref-breiman_random_2001}{}%
Breiman, Leo. 2001a. {``Random Forests.''} \emph{Machine Learning} 45
(1): 5--32.

\leavevmode\hypertarget{ref-breiman_statistical_2001}{}%
---------. 2001b. {``Statistical Modeling: {The} Two Cultures (with
Comments and a Rejoinder by the Author).''} \emph{Statistical Science}
16 (3): 199--231.

\leavevmode\hypertarget{ref-buja_grand_1986}{}%
Buja, Andreas, and Daniel Asimov. 1986. {``Grand {Tour} {Methods}: {An}
{Outline}.''} In \emph{Proceedings of the {Seventeenth} {Symposium} on
the {Interface} of {Computer} {Sciences} and {Statistics} on {Computer}
{Science} and {Statistics}}, 63--67. New York, NY, USA: Elsevier
North-Holland, Inc. \url{http://dl.acm.org/citation.cfm?id=26036.26046}.

\leavevmode\hypertarget{ref-chang_shiny_2021}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke,
Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara
Borges. 2021. \emph{Shiny: Web Application Framework for r}.
\url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-chen_xgboost_2021}{}%
Chen, Tianqi, Tong He, Michael Benesty, Vadim Khotilovich, Yuan Tang,
Hyunsu Cho, Kailong Chen, et al. 2021. {``Xgboost: {Extreme} {Gradient}
{Boosting}.''} \url{https://CRAN.R-project.org/package=xgboost}.

\leavevmode\hypertarget{ref-cook_manual_1997}{}%
Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
Computational and Graphical Statistics} 6 (4): 464--80.
\url{https://doi.org/10.2307/1390747}.

\leavevmode\hypertarget{ref-cook_grand_2008}{}%
Cook, Dianne, Andreas Buja, Eun-Kyung Lee, and Hadley Wickham. 2008.
{``Grand {Tours}, {Projection} {Pursuit} {Guided} {Tours}, and {Manual}
{Controls}.''} In \emph{Handbook of {Data} {Visualization}}, 295--314.
Berlin, Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-540-33037-0_13}.

\leavevmode\hypertarget{ref-da_silva_projection_2021}{}%
da Silva, Natalia, Dianne Cook, and Eun-Kyung Lee. 2021. {``A
{Projection} {Pursuit} {Forest} {Algorithm} for {Supervised}
{Classification}.''} \emph{Journal of Computational and Graphical
Statistics}, 1--21.

\leavevmode\hypertarget{ref-dastin_amazon_2018}{}%
Dastin, Jeffrey. 2018. {``Amazon Scraps Secret {AI} Recruiting Tool That
Showed Bias Against Women.''} \emph{Reuters}, October.
\url{https://www.reuters.com/article/us-amazon-com-jobs-automation-insight-idUSKCN1MK08G}.

\leavevmode\hypertarget{ref-diaz_addressing_2018}{}%
Díaz, Mark, Isaac Johnson, Amanda Lazar, Anne Marie Piper, and Darren
Gergle. 2018. {``Addressing Age-Related Bias in Sentiment Analysis.''}
In \emph{Proceedings of the 2018 Chi Conference on Human Factors in
Computing Systems}, 1--14.

\leavevmode\hypertarget{ref-duffy_apple_2019}{}%
Duffy, Claire. 2019. {``Apple Co-Founder {Steve} {Wozniak} Says {Apple}
{Card} Discriminated Against His Wife.''} \emph{CNN}, November.
\url{https://www.cnn.com/2019/11/10/business/goldman-sachs-apple-card-discrimination/index.html}.

\leavevmode\hypertarget{ref-espadoto_toward_2021}{}%
Espadoto, Mateus, Rafael M. Martins, Andreas Kerren, Nina S. T. Hirata,
and Alexandru C. Telea. 2021. {``Toward a {Quantitative} {Survey} of
{Dimension} {Reduction} {Techniques}.''} \emph{IEEE Transactions on
Visualization and Computer Graphics} 27 (3): 2153--73.
\url{https://doi.org/10.1109/TVCG.2019.2944182}.

\leavevmode\hypertarget{ref-gabriel_biplot_1971}{}%
Gabriel, Karl Ruben. 1971. {``The Biplot Graphic Display of Matrices
with Application to Principal Component Analysis.''} \emph{Biometrika}
58 (3): 453--67.

\leavevmode\hypertarget{ref-gorman_ecological_2014}{}%
Gorman, Kristen B., Tony D. Williams, and William R. Fraser. 2014.
{``Ecological Sexual Dimorphism and Environmental Variability Within a
Community of {Antarctic} Penguins (Genus {Pygoscelis}).''} \emph{PloS
One} 9 (3): e90081.

\leavevmode\hypertarget{ref-gosiewska_ibreakdown_2019}{}%
Gosiewska, Alicja, and Przemyslaw Biecek. 2019. {``{IBreakDown}:
{Uncertainty} of Model Explanations for Non-Additive Predictive
Models.''} \emph{arXiv Preprint arXiv:1903.11420}.

\leavevmode\hypertarget{ref-greenwell_fastshap_2020}{}%
Greenwell, Brandon. 2020. \emph{Fastshap: {Fast} {Approximate} {Shapley}
{Values}}. \url{https://CRAN.R-project.org/package=fastshap}.

\leavevmode\hypertarget{ref-greenwell_gbm_2020}{}%
Greenwell, Brandon, Bradley Boehmke, Jay Cunningham, and G. B. M.
Developers. 2020. {``Gbm: {Generalized} {Boosted} {Regression}
{Models}.''} \url{https://CRAN.R-project.org/package=gbm}.

\leavevmode\hypertarget{ref-horst_palmerpenguins_2020}{}%
Horst, Allison Marie, Alison Presmanes Hill, and Kristen B. Gorman.
2020. {``Palmerpenguins: {Palmer} {Archipelago} ({Antarctica}) Penguin
Data.''} \url{https://allisonhorst.github.io/palmerpenguins/}.

\leavevmode\hypertarget{ref-karwowski_international_2006}{}%
Karwowski, Waldemar. 2006. \emph{International {Encyclopedia} of
{Ergonomics} and {Human} {Factors}, -3 {Volume} {Set}}. CRC Press.

\leavevmode\hypertarget{ref-kodiyan_overview_2019}{}%
Kodiyan, Akhil Alfons. 2019. {``An Overview of Ethical Issues in Using
{AI} Systems in Hiring with a Case Study of {Amazon}'s {AI} Based Hiring
Tool.''} \emph{Researchgate Preprint}.

\leavevmode\hypertarget{ref-kominsarczyk_treeshap_2021}{}%
Kominsarczyk, Konrad, Pawel Kozminski, Szymon Maksymiuk, and Przemyslaw
Biecek. 2021. {``Treeshap.''} Model Oriented.
\url{https://github.com/ModelOriented/treeshap}.

\leavevmode\hypertarget{ref-larson_how_2016}{}%
Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin. 2016.
{``How {We} {Analyzed} the {COMPAS} {Recidivism} {Algorithm}.''}
\emph{ProPublica}, May.
\url{https://www.propublica.org/article/how-we-analyzed-the-compas-recidivism-algorithm?token=RPR1E2qtzJltfJ0tS-gB_41kmfoWZAu4}.

\leavevmode\hypertarget{ref-lee_state_2021}{}%
Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Nicholas
Spyrison, Earo Wang, and H. Sherry Zhang. 2021. {``The State-of-the-Art
on Tours for Dynamic Visualization of High-Dimensional Data.''}
\emph{WIREs Computational Statistics} n/a (n/a): e1573.
\url{https://doi.org/10.1002/wics.1573}.

\leavevmode\hypertarget{ref-lee_pptree_2013}{}%
Lee, Yoon Dong, Dianne Cook, Ji-won Park, and Eun-Kyung Lee. 2013.
{``{PPtree}: {Projection} Pursuit Classification Tree.''}
\emph{Electronic Journal of Statistics} 7: 1369--86.

\leavevmode\hypertarget{ref-leone_fifa_2020}{}%
Leone, Stefano. 2020. {``{FIFA} 20 Complete Player Dataset.''}
\url{https://kaggle.com/stefanoleone992/fifa-20-complete-player-dataset}.

\leavevmode\hypertarget{ref-liaw_classification_2002}{}%
Liaw, Andy, and Matthew Wiener. 2002. {``Classification and Regression
by {randomForest}.''} \emph{R News} 2 (3): 18--22.

\leavevmode\hypertarget{ref-lundberg_consistent_2018}{}%
Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2018. {``Consistent
Individualized Feature Attribution for Tree Ensembles.''} \emph{arXiv
Preprint arXiv:1802.03888}.

\leavevmode\hypertarget{ref-lundberg_unified_2017}{}%
Lundberg, Scott M., and Su-In Lee. 2017. {``A Unified Approach to
Interpreting Model Predictions.''} In \emph{Proceedings of the 31st
International Conference on Neural Information Processing Systems},
4768--77.

\leavevmode\hypertarget{ref-molnar_interpretable_2020}{}%
Molnar, Christoph. 2020. \emph{Interpretable Machine Learning}. Lulu.
com.
\href{https://christophm.github.io/interpretable-ml-book/}{christophm.github.io/interpretable-ml-book/}.

\leavevmode\hypertarget{ref-oneil_weapons_2016}{}%
O'Neil, Cathy. 2016. \emph{Weapons of Math Destruction: {How} Big Data
Increases Inequality and Threatens Democracy}. Crown.

\leavevmode\hypertarget{ref-ocagne_coordonnees_1885}{}%
Ocagne, Maurice d'. 1885. \emph{Coordonnées Parallèles Et Axiales.
{Méthode} de Transformation géométrique Et Procédé Nouveau de Calcul
Graphique déduits de La Considération Des Coordonnées Parallèles, Par
{Maurice} d'{Ocagne}, ...} Paris: Gauthier-Villars.

\leavevmode\hypertarget{ref-ribeiro_why_2016}{}%
Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin. 2016. {``"{Why}
{Should} {I} {Trust} {You}?": {Explaining} the {Predictions} of {Any}
{Classifier}.''} In \emph{Proceedings of the 22nd {ACM} {SIGKDD}
{International} {Conference} on {Knowledge} {Discovery} and {Data}
{Mining}}, 1135--44. {KDD} '16. New York, NY, USA: Association for
Computing Machinery. \url{https://doi.org/10.1145/2939672.2939778}.

\leavevmode\hypertarget{ref-roberts_state_2007}{}%
Roberts, Jonathan C. 2007. {``State of the Art: {Coordinated} \&
Multiple Views in Exploratory Visualization.''} In \emph{Fifth
International Conference on Coordinated and Multiple Views in
Exploratory Visualization ({CMV} 2007)}, 61--71. IEEE.

\leavevmode\hypertarget{ref-shapley_value_1953}{}%
Shapley, Lloyd S. 1953. \emph{A Value for n-Person Games}. Princeton
University Press.

\leavevmode\hypertarget{ref-shi_lightgbm_2022}{}%
Shi, Yu, Guolin Ke, Damien Soukhavong, James Lamb, Qi Meng, Thomas
Finley, Taifeng Wang, et al. 2022. {``Lightgbm: {Light} {Gradient}
{Boosting} {Machine}.''}
\url{https://CRAN.R-project.org/package=lightgbm}.

\leavevmode\hypertarget{ref-shmueli_explain_2010}{}%
Shmueli, Galit. 2010. {``To Explain or to Predict?''} \emph{Statistical
Science} 25 (3): 289--310.

\leavevmode\hypertarget{ref-shrikumar_learning_2017}{}%
Shrikumar, Avanti, Peyton Greenside, and Anshul Kundaje. 2017.
{``Learning Important Features Through Propagating Activation
Differences.''} In \emph{International {Conference} on {Machine}
{Learning}}, 3145--53. PMLR.

\leavevmode\hypertarget{ref-shrikumar_not_2016}{}%
Shrikumar, Avanti, Peyton Greenside, Anna Shcherbina, and Anshul
Kundaje. 2016. {``Not Just a Black Box: {Learning} Important Features
Through Propagating Activation Differences.''} \emph{arXiv Preprint
arXiv:1605.01713}.

\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data}
{Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC.
\url{https://plotly-r.com}.

\leavevmode\hypertarget{ref-simonyan_deep_2014}{}%
Simonyan, Karen, Andrea Vedaldi, and Andrew Zisserman. 2014. {``Deep
Inside Convolutional Networks: {Visualising} Image Classification Models
and Saliency Maps.''} In \emph{In {Workshop} at {International}
{Conference} on {Learning} {Representations}}. Citeseer.

\leavevmode\hypertarget{ref-spyrison_spinifex_2020}{}%
Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
\url{https://doi.org/10.32614/RJ-2020-027}.

\leavevmode\hypertarget{ref-strumbelj_efficient_2010}{}%
Strumbelj, Erik, and Igor Kononenko. 2010. {``An Efficient Explanation
of Individual Classifications Using Game Theory.''} \emph{The Journal of
Machine Learning Research} 11: 1--18.

\leavevmode\hypertarget{ref-unwin_ensemble_2018}{}%
Unwin, Antony, and Pedro Valero-Mora. 2018. {``Ensemble {Graphics}.''}
\emph{Journal of Computational and Graphical Statistics} 27 (1):
157--65. \url{https://doi.org/10.1080/10618600.2017.1383264}.

\leavevmode\hypertarget{ref-vanni_textual_2018}{}%
Vanni, Laurent, Mélanie Ducoffe, Carlos Aguilar, Frédéric Precioso, and
Damon Mayaffre. 2018. {``Textual {Deconvolution} {Saliency} ({TDS}): A
Deep Tool Box for Linguistic Analysis.''} In \emph{Proceedings of the
56th {Annual} {Meeting} of the {Association} for {Computational}
{Linguistics} ({Volume} 1: {Long} {Papers})}, 548--57.

\leavevmode\hypertarget{ref-wickham_ggplot2_2016}{}%
Wickham, Hadley. 2016. \emph{Ggplot2: {Elegant} {Graphics} for {Data}
{Analysis}}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-wickham_visualizing_2015}{}%
Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
Statistical Models: {Removing} the Blindfold.''} \emph{Statistical
Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
\url{https://doi.org/10.1002/sam.11271}.

\leavevmode\hypertarget{ref-wright_ranger_2017}{}%
Wright, Marvin N., and Andreas Ziegler. 2017. {``Ranger: {A} {Fast}
{Implementation} of {Random} {Forests} for {High} {Dimensional} {Data}
in {C}++ and {R}.''} \emph{Journal of Statistical Software} 77 (1):
1--17. \url{https://doi.org/10.18637/jss.v077.i01}.

\end{CSLReferences}

\end{document}