### 2022-03-14 z_sty_reprex/sty_reprex.rmd

---
title: "title"
author: 
  - author
  - author2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header:
        #- ../template/jmlr2e.sty ## Doesn't work
        #- tmp_jmlr.tex ## Doesn't work
        - mystyles.sty ## Works
---

A minimal reprex following https://stackoverflow.com/questions/56174593/how-to-add-sty-file-to-rmarkdown-pdf-output.

# Section 1

Hello world