### 2022-03-14 z_sty_reprex/sty_reprex.rmd
partially working .sty (z_sty_reprex/)

---
title: "title"
author: 
  - author
  - author2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header:
        - jmlr2e.sty ## Doesn't work
        #- mystyles.sty ## Works (trivial \usepackage{})
---

A minimal reprex following https://stackoverflow.com/questions/56174593/how-to-add-sty-file-to-rmarkdown-pdf-output.

# Section 1

Hello world

# Attempts and notes

- Trivial sty files work (mystyles.sty)
- Trivial tex file didn't work.
- jmlr2e.sty doesn't work wholesale, hence:
- Changing to *.tex didn't work.
- Trying to pare back and to isolate the issue:
    - adding \makeatletter and \makeatother to change the order of operations handling @ in marco assignment. _partial success!!_
    - uncommenting .sty file, keeping an eye out for @ in macronames. 
    - \makeatletter and \makeatother at top and bottom is back to original error... (! LaTeX Error: Missing \begin{document}.); try changing back and forth