### 2022-03-15 z_sty_reprex/sty_reprex.rmd
sty reprex progress atm.

---
title: "title"
author:
  - author
  - author2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header:
        - jmlr2e.sty          ## Applying comments and changes here
        #- jmlr2eoriginal.sty ## Not working; full original .sty
        #- mystyles.sty       ## Works (with trivial \usepackage{})
---

A minimal reprex following https://stackoverflow.com/questions/56174593/how-to-add-sty-file-to-rmarkdown-pdf-output.

# Section title

Hello world

# Attempts and notes

Primary errors:

- original error: `! LaTeX Error: Missing \begin{document}.`
- alternative error: `! You can't use \spacefactor in vertical mode.`

Attempt notes:

- Trivial sty files work (mystyles.sty)
- Trivial tex file didn't work.
- jmlr2e.sty doesn't work wholesale, hence:
- Changing to *.tex didn't work.
- Trying to pare back and to isolate the issue:
    - adding \makeatletter and \makeatother to change the order of operations handling @ in marco assignment. _partial success!!_
    - uncommenting .sty file, keeping an eye out for @ in macronames.
    - \makeatletter and \makeatother at top and bottom is back to original error... (); try changing back and forth