### 2022-03-16 jmlr.tex

\documentclass[11pt,twoside]{article}

% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
%
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf

% Available options for package jmlr2e are:
%
%   - abbrvbib : use abbrvnat for the bibliography style
%   - nohyperref : do not load the hyperref package
%   - preprint : remove JMLR specific information from the template,
%         useful for example for posting to preprint servers.
%
% Example of using the package with custom options:
%
% \usepackage[abbrvbib, preprint]{jmlr2e}

\usepackage[]{jmlr2e}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Definitions of handy macros can go here

%% Any special functions or other packages can be loaded here.

% Heading arguments are {volume}{year}{pages}{date submitted}{date published}{paper id}{author-full-names}

\jmlrheading{}{2022}{}{03/22}{--}{}{%
Fan Cheng, Anastasios Panagiotelis and Rob J Hyndman}

% Short headings should be running head and authors last names

\ShortHeadings{Shorter title}{Cheng, Panagiotelis and Hyndman}
\firstpageno{1}

\begin{document}
% Add back JMLR proof environment which is over-written by amsthm
\def\proof{\par\noindent{\bf Proof\ }}
\def\endproof{\hfill\BlackBox\\[2mm]}

\title{Title}

\author{%
\name Fan Cheng \email \href{mailto:Fan.Cheng@monash.edu}{\nolinkurl{Fan.Cheng@monash.edu}} \\\addr Monash University\newline Australia \AND
%
\name Anastasios Panagiotelis \email \href{mailto:Anastasios.Panagiotelis@sydney.edu.au}{\nolinkurl{Anastasios.Panagiotelis@sydney.edu.au}} \\\addr University of Sydney\newline Australia \AND
%
\name Rob J Hyndman \email \href{mailto:Rob.Hyndman@monash.edu}{\nolinkurl{Rob.Hyndman@monash.edu}} \\\addr Monash University\newline Australia \AND
}

\editor{}

\maketitle

\begin{abstract}
% Put your abstract here. Abstracts are limited to 200 words for
% regular articles and 100 words for Letters to the Editor. Please no
% personal pronouns, also please do not use the words ``new'' and/or
% ``novel'' in the abstract. An article usually includes an abstract, a
% concise summary of the work covered at length in the main body of the
% article.
Blah blah blah
\end{abstract}

\begin{keywords}
dimension reduction, divergence estimation, Hellinger distance, anomaly detection, k-d trees
\end{keywords}

%  Body of the article
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Blah blah. You can refer to papers \citep{Witten2011} or cite them directly as in \citet{AriasCastro2020}.

Add the following to the yaml header if you want to remove the JMLR details.

\begin{verbatim}
jmlroption: preprint
\end{verbatim}

% Acknowledgements should go at the end, before appendices and references

\acks{This research was supported in part by the\ldots{}}

% Appendices to go in include-after

\vskip 0.2in
\bibliography{references.bib}

\end{document}