### 2022-03-16 template_rjh/jmlr.Rmd

---
title: "Title"
shorttitle: "Shorter title"
author:
- familyname: Cheng
  othernames: Fan
  address: Monash University\newline Australia
  email: Fan.Cheng@monash.edu
  correspondingauthor: true
- familyname: Panagiotelis
  othernames: Anastasios
  address: University of Sydney\newline Australia
  email: Anastasios.Panagiotelis@sydney.edu.au
- familyname: Hyndman
  othernames: Rob J
  address: Monash University\newline Australia
  email: Rob.Hyndman@monash.edu
year: 2022
date_submitted: "03/22"
abstract: "Blah blah blah"
keywords: "dimension reduction, divergence estimation, Hellinger distance, anomaly detection, k-d trees"
acknowledgment: "This research was supported in part by the..."
bibliography: references.bib
classoption: 11pt,twoside
output:
  bookdown::pdf_document2:
    template: jmlr_template.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: natbib
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE
  # include = FALSE
)
library(tidyverse)
```

\newpage

# Introduction

Blah blah. You can refer to papers [@Witten2011] or cite them directly as in @AriasCastro2020.

Add the following to the yaml header if you want to remove the JMLR details.

```
jmlroption: preprint
```