### 2023-05-21 jss/cheem_paper.tex

\documentclass[
]{jss}

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

\usepackage[utf8]{inputenc}

\author{
Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
University \And Przemyslaw
Biecek~\orcidlink{0000-0001-8423-1823}\\Warsaw University of\\
Technology and University
}
\title{Exploring Local Explanations of Nonlinear Models Using Animated
Linear Projections}

\Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}
\Plaintitle{Exploring Local Explanations of Nonlinear Models Using
Animated Linear Projections}
\Shorttitle{Exploring Local Explanations}

\Abstract{
The increased predictive power of machine learning models comes at the
cost of increased complexity and loss of interpretability, particularly
in comparison to parametric statistical models. This trade-off has led
to the emergence of eXplainable AI (XAI) which provides methods, such as
local explanations (LEs) and local variable attributions (LVAs), to shed
light on how a model use predictors to arrive at a prediction. These
provide a point estimate of the linear variable importance in the
vicinity of a single observation. However, LVAs tend not to effectively
handle association between predictors. To understand how the interaction
between predictors affects the variable importance estimate, we can
convert LVAs into linear projections and use the radial tour. This is
also useful for learning how a model has made a mistake, or the effect
of outliers, or clustering of observations. The approach is llustrated
with examples from categorical (penguin species, chocolate types) and
quantitative (soccer/football salaries, house prices) response models.
The methods are implemented in the R package cheem, available on CRAN.
}

\Keywords{explainable artificial intelligence, nonlinear model
interpretability, visual analytics, local explanations, grand
tour, radial tour}
\Plainkeywords{explainable artificial intelligence, nonlinear model
interpretability, visual analytics, local explanations, grand
tour, radial tour}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Nicholas Spyrison\\
    Monash University\\
    Dept of Human Centred Computing\\
Faculty of Information Technology\\
  E-mail: \email{spyrison@gmail.com}\\
  URL: \url{https://nspyrison.netlify.app}\\~\\
      Dianne Cook\\
    Monash University\\
    Dept of Econometrics and Business Statistics\\

    }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{amsmath}

\begin{document}

\hypertarget{sec:intro}{%
\section{Introduction}\label{sec:intro}}

There are different reasons and purposes for fitting a model. According
to the taxonomies of \citet{breiman_statistical_2001} and
\citet{shmueli_explain_2010}, it can be useful to group models into two
types: explanatory and predictive. Explanatory modeling is used for
inferential purposes, while predictive modeling focuses solely on the
performance of an objective function. The intended use of the model has
important implications for its selection and development.
Interpretability is critical in explanatory modeling to draw meaningful
inferential conclusions, such as which variables most contribute to a
prediction or whether some observations are less well fit.
Interpretability becomes more difficult when the model is nonlinear.
Nonlinear models occur in statistical models with polynomial or
interaction terms between quantitative predictors, and almost all
computational models such as random forests, support-vector machines, or
neural networks
\citep[e.g.][]{breiman_random_2001, boser_training_1992, anderson_introduction_1995}.

In linear models interpretation of the importance of variables is
relatively straight forward, one adjusts for the covariance of multiple
variables when examining the relationship with the response. The
interpretation is valid for the full domain of the predictors. In
nonlinear models one needs to consider the model in small neighborhoods
of the domain to make any assessment of variable importance. Even though
this is difficult, it is especially important to interpret model fits as
we become more dependent on nonlinear models for routine aspects of life
to avoid issues described in \citet{stahl-ethics}. Understanding how
nonlinear models behave when usage extrapolates outside the domain of
predictors, either in sub-spaces where few samples were provided in the
training set, or extending outside the domain. It is especially
important because nonlinear models can vary wildly and predictions can
be dramatically wrong in these areas.

Explainable Artificial Intelligence (XAI) is an emerging field of
research focused on methods for the interpreting of models
\citep{adadi_peeking_2018, arrieta_explainable_2020}. A class of
techniques, called \emph{local explanations} (LEs), provide methods to
approximate linear variable importance, called local variable
attributions (LVAs), at the location of each observation or the
predictions at a specific point in the data domain. Because these are
point-specific, it is challenging to comprehensively visualize them to
understand a model. There are common approaches for visualizing
high-dimensional data as a whole, but what is needed are new approaches
for viewing these individual LVAs relative to the whole.

For multivariate data visualization, a \emph{tour}
\citep{asimov_grand_1985, buja_grand_1986, lee_state_2021} of linear
data projections onto a lower-dimensional space, could be an element of
XAI, complementing LVAs. Applying tours to model interpretation is
recommended in \citet{wickham_visualizing_2015} primarily to examine the
fitted model in the space of the data. \citet{cook_interactive_2007}
describe the use of tours for exploring classification boundaries and
model diagnostics
\citep{Caragea2008, lee_pptree_2013, da_silva_projection_2021}. There
are various types of tours. In a \emph{manual} or radial tour
\citep{cook_manual_1997, spyrison_spinifex_2020}, the path of linear
projections is defined by changing the contribution of a selected
variable. We propose to use this to scrutinize the LVAs. This approach
could be considered to be a counter-factual, what-if analysis, such as
\emph{ceteris paribus} (``other things held constant'') profiles
\citep{biecek_ceterisparibus_2020}.

The remainder of this paper is organized as follows. Section
\ref{sec:explanations} covers the background of the LEs and the
traditional visuals produced. Section \ref{sec:tour} explains the tours
and particularly the radial manual tour. Section \ref{sec:cheemviewer}
discusses the visual layout in the graphical user interface and how it
facilitates analysis, data pre-processing, and package infrastructure.
Illustrations are provided in Section \ref{sec:casestudies} for a range
of supervised learning tasks with categorical and quantitative response
variables. These show how the LVAs can be used to get an overview of the
model's use of predictors, and to investigate errors in the model
predictions. Section \ref{sec:cheemdiscussion} concludes with a summary
of the insights gained. The methods are implemented in the \textbf{R}
package \textbf{cheem}.

\hypertarget{sec:explanations}{%
\section{Local Explanations}\label{sec:explanations}}

LVAs shed light on machine learning model fits by estimating linear
variable importance in the vicinity of a single observation. There are
many approaches for calculating LVAs. A comprehensive summary of the
taxonomy of currently available methods is provided in Figure 6 by
\citet{arrieta_explainable_2020}. It includes a large number of
model-specific explanations such as deepLIFT
\citep{shrikumar_not_2016, shrikumar_learning_2017}, a popular recursive
method for estimating importance in neural networks. There are fewer
model-agnostic methods, of which LIME \citep{ribeiro_why_2016} and
SHaply Additive exPlanations (SHAP) \citep{lundberg_unified_2017}, are
popular.

These observation-level explanations are used in various ways depending
on the data. In image classification, where pixels correspond to
predictors, saliency maps overlay or offset a heatmap to indicate
important pixels \citep{simonyan_deep_2014}. For example, pixels
corresponding to snow may be highlighted as important contributors when
distinguishing if a picture contains a coyote or husky. In text
analysis, word-level contextual sentiment analysis highlights the
sentiment and magnitude of influential words \citep{vanni_textual_2018}.
In the case of numeric regression, they are used to explain additive
contributions of variables from the model intercept to the observation's
prediction \citep{ribeiro_why_2016}.

We will be focusing on SHAP values in this this paper, but the approach
is applicable for any method used to calculate the LVAs. SHAP calculates
the variable contributions of one observation by examining the effect of
other variables on the predictions. The term ``SHAP'' refers to
\citet{shapley_value_1953}'s method to evaluate an individual's
contribution in cooperative games by assessing this player's performance
in the presence or absence of other players.
\citet{strumbelj_efficient_2010} introduced SHAP for LEs in machine
learning models. Variable importance can depend on the sequence in which
variables are entered into the model fitting process, thus for any
sequence we get a set of variable contribution values for a single
observation. These values will add up to the difference between the
fitted value for the observation, and the average fitted value for all
observations. Using all possible sequence, or permutation, gives
multiple values for each variable, which are averaged to get the SHAP
value for an observation. It can be helpful to standardize variables
prior to computing SHAP values, if they have been measured on different
scales.

The approach is related to partial dependence plots (see for example,
chapter 8 of \citet{molnar2022}), used to explain the effect of a
variable by predicting the response for a range of values on this
variable after fixing the value of all other variables to their mean.
Though partial dependence plots are a global approximation of the
variable importance, while SHAP is specific to one observation.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}

}

\caption[Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors]{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) shows breakdown plots of three sequences of the variables. The sequence of the variables impacts the magnitude of their attribution. Panel (b) shows the distribution of attribution for each variable across 25 sequences of predictors, with the mean displayed as a dot for each player. Reaction skills are important for both players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
\end{figure}
\end{CodeChunk}

We use 2020 season FIFA data \citep{leone_fifa_2020} to illustrate SHAP
following the procedures described in \citet{biecek_explanatory_2021}.
There are 5000 observations of nine predictor variables measuring
players' skills and one response variable, wages (in euros). A random
forest model is fit regressing players' wages on the skill variables. In
this illustration in Figure \ref{fig:shapdistrbd} the SHAP values are
compared for a star offensive player (L. Messi) and a prominent
defensive player (V. van Dijk). We are interested in knowing how the
skill variables locally contribute to the wage prediction of each
player. A difference in the attribution of the variable importance
across the two positions of the players can be expected. This would be
interpreted as how a player's salary depends on which combination of
skills. Panel (a) is a version of a breakdown plot
\citep{gosiewska_ibreakdown_2019} where just three sequences of
variables are shown, for two observations. A breakdown plot shows the
absolute values of the variable attribution for an observation, usually
sorted from highest value to the lowest. There is no scale on the
horizontal axis here because values are considered relative to each
other. Here we can see how the variable contribution can change
depending on sequence, relative to both players. (Note that the order of
the variables is different in each plot, because they have been sorted
by biggest average contribution across both players.) For all sequences,
and for both players \texttt{reaction} has the strongest contribution,
with perhaps more importance for the defensive player. Then it differs
by player: for Messi \texttt{offense} and \texttt{movement} have the
strongest contributions, and for van Dijk it is \texttt{defense} and
\texttt{power}, regardless of the variable sequence.

Panel (b) shows the differences in the player's median values (large
dots) for 25 such sequences (tick marks). We can see that the wage
predictions for the two players come from different combinations of
skills sets, as might be expected for players who's value on the team
depends on their offensive or defensive prowess. It is also interesting
to see from the distribution of values across the different sequence of
variables, that there is some multimodality. For example, look at the
SHAP values for \texttt{reaction} for Messi, and in some sequences
reaction has a much lower contribution than others. This suggests that
other variables (\texttt{offense}, \texttt{movement} probably) can
substitute for \texttt{reaction} in the wage prediction.

This can also be considered similar to examining the coefficients from
all subsets regression, as described in
\citet{wickham_visualizing_2015}. Various models that are similarly good
might use different combinations of the variables. Examining the
coefficients from multiple models helps to understand the relative
importance of each variable in the context of all other variables. This
is similar to the approach here with SHAP values, that by examining the
variation in values across different permutations of variables, we can
gain more understanding of the relationship between the response and
predictors.

For the application, we use \emph{tree SHAP}, a variant of SHAP that
enjoys a lower computational complexity
\citep{lundberg_consistent_2018}. Instead of aggregating over sequences
of the variables, tree SHAP calculates observation-level variable
importance by exploring the structure of the decision trees. Tree SHAP
is only compatible with tree-based models. so random forests are used
for illustration.

There are numerous R packages currently available which provide
functions for computing SHAP values, including \texttt{fastshap}
\citep{fastshap}, \texttt{kernelshap} \citep{kernelshap}, \texttt{shapr}
\citep{shapr}, \texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
\citep{PPtreeregViz}, and the package \texttt{DALEX} has many resources
\citep{biecek_dalex_2018}. There are many more packages only available
through Github, like \texttt{treeshap}
{[}kominsarczyk\_treeshap\_2021{]} that is used for this work.
\citet{molnar2022} provides good explanations of the different methods
and how to apply them to different models.

\hypertarget{sec:tour}{%
\section{Tours and the Radial Tour}\label{sec:tour}}

A \emph{tour} enables the viewing of high-dimensional data by animating
many linear projections with small incremental changes. It is achieved
by following a path of linear projections (bases) of high-dimensional
space. One key variable of the tour is the object permanence of the data
points; one can track the relative change of observations in time and
gain information about the relationships between points across multiple
variables. There are various types of tours that are distinguished by
how the paths are generated \citep{lee_state_2021, cook_grand_2008}.

The manual tour \citep{cook_manual_1997} defines its path by changing a
selected variable's contribution to a basis to allow the variable to
contribute more or less to the projection. The requirement constrains
the contribution of all other variables that a basis needs to be
orthonormal (column correspond to vectors, with unit length, and
orthogonal to each other). The manual tour is primarily used to assess
the importance of a variable to structure visible in a projection. It
also lends itself to pre-computation queued in advance or computed
on-the-fly for human-in-the-loop analysis
\citep{karwowski_international_2006}.

A version of the manual tour called a \emph{radial tour} is implemented
in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
In a radial tour, the selected variable can change its magnitude of
contribution but not its angle; it must move along the direction of its
original contribution. The implementation allows for pre-computation and
interactive re-calculation to focus on a different variable.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.99\linewidth]{./figures/radial_tour}

}

\caption{The radial tour allows the user to remove a variable from a projection, to examine the importance of this variable to structure in the plot. Here we have a 1D projection of the penguins data displayed as a density plot. The line segments on the bottom correspond to the coefficients of the variables making up the projection. The structure in the plot is bimodality (left), and the importance of the variable \textsf{bd} is being explored. As this variable contribution is reduced in the plot (middle, right) we can see that the bimodality decreases. Thus \textsf{bd} is an important variable contributing to the bimodal structure.}\label{fig:radialtour}
\end{figure}
\end{CodeChunk}

\hypertarget{sec:cheemviewer}{%
\section{The Cheem Viewer}\label{sec:cheemviewer}}

To explore the LVAs, coordinated views \citep{roberts_state_2007}
\citep[also known as ensemble graphics,][]{unwin_ensemble_2018} are
provided in the \emph{cheem viewer} application. There are two primary
plots: the \textbf{global view} to give the context of all of the SHAP
values and the \textbf{radial tour view} to explore the LVAs with
user-controlled rotation. There are numerous user inputs, including
variable selection for the radial tour and observation selection for
making comparisons. There are different plots used for the categorical
and quantitative responses. Figures \ref{fig:classificationcase} and
\ref{fig:regressioncase} are screenshots showing the cheem viewer for
the two primary tasks: classification (categorical response) and
regression (quantitative response).

\hypertarget{global-view}{%
\subsection{Global View}\label{global-view}}

The global view provides context for all observations and facilitates
the exploration of the separability of the data and attribution spaces.
The attribution space refers to the SHAP values for each observation.
These spaces both have dimensionality \(n \times p\), where \(n\) is the
number of observations and \(p\) is the number of variables.

The visualization is composed of the first two principal components of
the data (left) and the attribution (middle) spaces. These single 2D
projections will not reveal all of the structure of higher-dimensional
space, but they are helpful visual summaries. In addition, a plot of the
observed against predicted response values is also provided (Figures
\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
identify observations poorly predicted by the model. For classification
tasks, color indicates the predicted class, and misclassified
observations are circled in red. Linked brushing between the plots is
provided, and a tabular display of selected points helps to facilitate
exploration of the spaces and the model (shown in Figures
\ref{fig:regressioncase}d).

While the comparison of these spaces is interesting, the primary purpose
of the global view is to enable the selection of particular observations
to explore in detail. We have designed it to enable a comparison between
an observation that is interesting is some way, perhaps misclassified,
or poorly predicted, relative to an observation with similar predictor
values but more expected prediction. For brevity, we call the
interesting observation the primary investigation (PI) and the other is
the comparison investigation (CI). These observations are highlighted as
asterisk and \(\times\), respectively.

\hypertarget{radial-tour}{%
\subsection{Radial Tour}\label{radial-tour}}

There are two plots in this part of the interface. The first (Figures
\ref{fig:classificationcase}e and \ref{fig:regressioncase}e) is a
display of the SHAP values for all observations. This will generally
give the global view of variables important for the fit as a whole, but
it will also highlight observations which have different patterns. The
second plot is the radial tour, which for classification is a density
plot of a 1D projection (Figure \ref{fig:classificationcase}f), and for
regression are scatterplots of the observed response values, and
residuals, against a 1D projection (Figure \ref{fig:regressioncase}f).

The LVAs for all observations are normalized (sum of squares equals 1),
and thus, the relative importance of variables can be compared across
all observations. These are depicted as a vertical parallel coordinate
plot \citep{ocagne_coordonnees_1885}. (The SHAP values of the PI and CI
are shown as dashed and dotted lines, respectively.) One should obtain a
sense of the overall importance of variables from this plot. The more
important variables will have larger values, and in the case of
classification tasks variables which have different magnitude for
different classes are more globally important. For example, Figure
\ref{fig:classificationcase}e suggests that \texttt{bl} is important for
distinguishing the green class from the other two. For regression, one
might generally observe which variables have low values for all
observations (not important). For example, \texttt{BMI} and \texttt{pwr}
in Figure \ref{fig:regressioncase}e, have a range of high and low values
(e.g., \texttt{off}, \texttt{def}), suggesting they are important for
some observations and not important for others.

A bar chart is overlaid to represent the projection shown in the radial
tour at right. It starts from the SHAP values of the PI, but if the user
changes the projection the length of these bars will reflect this
change. (The PI is interactively selected by clicking on a point in the
global view). By scaling the SHAP value it becomes an (attribution)
projection.

The attribution projection of the PI is the initial 1D basis in a radial
tour, displayed as a density plot for a categorical response (Figure
\ref{fig:classificationcase}f) and as scatterplots for a quantitative
response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
by vertical dashed and dotted lines, respectively. The radial tour
varies the contribution of the selected variable. This is viewed as an
animation of the projections from many intermediate bases. Doing so
tests the sensitivity of structure (class separation or strength of
relationship) to the variable's contribution. For classification, if the
separation between classes diminishes when the variable contribution is
reduced, this suggests that the variable is important for class
separation. For regression, if the relationship scatterplot weakens when
the variable contribution is reduced, indicating that the variable is
important for accurately predicting the response.

\hypertarget{classification-task}{%
\subsection{Classification Task}\label{classification-task}}

Selecting a misclassified observation as PI and a correctly classified
point nearby in data space as CI makes it easier to examine the
variables most responsible for the error. The global view (Figure
\ref{fig:classificationcase}c) displays the model confusion matrix. The
radial tour is 1D and displays as density where color indicates class.
An animation slider enables users to vary the contribution of variables
to explore the sensitivity of the separation to that variable.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/app_classification}

}

\caption[Overview of the cheem viewer for classification tasks (categorical response)]{Overview of the cheem viewer for classification tasks (categorical response). Global view inputs, (a), set the PI, CI, and color statistic. Global view, (b) PC1 by PC2 approximations of the data- and attribution-space. (c) prediction by observed $y$ (visual of the confusion matrix for classification tasks). Points are colored by predicted class, and red circles indicate misclassified observations. Radial tour inputs (d) select variables to include and which variable is changed in the tour. (e) shows a parallel coordinate display of the distribution of the variable attributions while bars depict contribution for the current basis. The black bar is the variable being changed in the radial tour. Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
\end{figure}
\end{CodeChunk}

\hypertarget{regression-task}{%
\subsection{Regression Task}\label{regression-task}}

Selecting an inaccurately predicted observation as PI and an accurately
predicted observation with similar variable values as CI is a helpful
way to understand how the model is failing or not. The global view
(Figure \ref{fig:regressioncase}a) shows a scatterplot of the observed
vs predicted values, which should exhibit a strong relationship if the
model is a good fit. The points can be colored by a statistic, residual,
a measure of outlyingness (log Mahalanobis distance), or correlation to
aid in understanding the structure identified in these spaces.

In the radial tour view, the observed response and the residuals
(vertical) are plotted against the attribution projection of the PI
(horizontal). The attribution projection can be interpreted similarly to
the predicted value from the global view plot. It represents a linear
combination of the variables, and a good fit would be indicated when
there is a strong relationship with the observed values. This can be
viewed as a local linear approximation if the fitted model is nonlinear.
As the contribution of a variable is varied, if the value of the PI does
not change much, it would indicate that the prediction for this
observation is NOT sensitive to that variable. Conversely, if the
predicted value varies substantially, the prediction is very sensitive
to that variable, suggesting that the variable is very important for the
PI's prediction.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}

}

\caption[Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables]{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables. Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values. Four selected points are highlighted in the PC spaces and tabularly displayed. Coloring on a statistic (c) highlights structure organized in the attribution space. Interactive tabular display (d) populates when observations are selected. Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis. Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
\end{figure}
\end{CodeChunk}

\hypertarget{interactive-variables}{%
\subsection{Interactive variables}\label{interactive-variables}}

The application has several reactive inputs that affect the data used,
aesthetic display, and tour manipulation. These reactive inputs make the
software flexible and extensible (Figure \ref{fig:classificationcase}a
\& d). The application also has more exploratory interactions to help
link points across displays, reveal structures found in different
spaces, and access the original data.

A tooltip displays observation number/name and classification
information while the cursor hovers over a point. Linked brushing allows
the selection of points (left click and drag) where those points will be
highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
The information corresponding to the selected points is populated on a
dynamic table (Figure \ref{fig:classificationcase}d). These interactions
aid exploration of the spaces and, finally, the identification of
primary and comparison observations.

\hypertarget{preprocessing}{%
\subsection{Preprocessing}\label{preprocessing}}

It is vital to mitigate the render time of visuals, especially when
users may want to iterate many explorations. All computational
operations should be prepared before run time. The work remaining when
an application is run solely reacts to inputs and rendering visuals and
tables. Below discusses the steps and details of the reprocessing.

\begin{itemize}
\tightlist
\item
  \textbf{Data:} predictors and response are unscaled complete numerical
  matrix. Most models and local explanations are scale-invariant. Keep
  normality assumptions of the model in mind.
\item
  \textbf{Model:} any model and compatible explanation could be explored
  with this method. Currently, random forest models are applied via the
  package \textbf{randomForest} \citep{liaw_classification_2002},
  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
  \(p/3\) for classification and regression, and minimum size of
  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
  classification and regression.
\item
  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
  observation using the package \textbf{treeshap}
  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
  each observation in the training data and not fit to fit variable
  interactions.
\item
  \textbf{Cheem viewer:} after the model and full explanation space are
  calculated, each variable is scaled by standard deviations away from
  the mean to achieve common support for visuals. Statistics for mapping
  to color are computed on the scaled spaces.
\end{itemize}

The time to preprocess the data will vary significantly with the
complexity of the model and the LE. For reference, the FIFA data
contained 5000 observations of nine explanatory variables took 2.5
seconds to fit a random forest model of modest hyperparameters.
Extracting the tree SHAP values of each observation took 270 seconds
total. PCA and statistics of the variables and attributions took 2.8
seconds. These run times were from a non-parallelized session on a
modern laptop, but suffice to say that most of the time will be spent on
the LVA. An increase in model complexity or data dimensionality will
quickly become an obstacle. Its reduced computational complexity makes
tree SHAP an excellent candidate to start. Alternatively, the package
\textbf{fastshap} claims extremely low run times, attributed to fewer
calls to the prediction function, partial implementation in C++, and
efficient use of logical subsetting, \citet{greenwell_fastshap_2020}.

\hypertarget{sec:casestudies}{%
\section{Case Studies}\label{sec:casestudies}}

To illustrate the cheem method it is applied to modern data sets, two
classification examples and then two of regression.

\hypertarget{palmer-penguin-species-classification}{%
\subsection{Palmer Penguin, Species
Classification}\label{palmer-penguin-species-classification}}

The Palmer penguins data
\citep{gorman_ecological_2014, horst_palmerpenguins_2020} was collected
on three species of penguins foraging near Palmer Station, Antarctica.
The data is publicly available to substitute for the overly-used iris
data and is quite similar in form. After removing incomplete
observations, there are 333 observations of four physical measurements,
bill length (\texttt{bl}), bill depth (\texttt{bd}), flipper length
(\texttt{fl}), and body mass (\texttt{bm}) for this illustration. A
random forest model was fit with species as the response variable.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}

}

\caption[Examining the SHAP values for a random forest model classifying Palmer penguin species]{Examining the SHAP values for a random forest model classifying Palmer penguin species. The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b). The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right). (The animation of the radial tour is at https://vimeo.com/666431172.)}\label{fig:casepenguins}
\end{figure}
\end{CodeChunk}

Figure \ref{fig:casepenguins} shows plots from the cheem viewer for
exploring the random forest model on the penguins data. Panel (a) shows
the global view, and panel (b) shows several 1D projections generated
with the radial tour. Penguin 243, a Gentoo (purple), is the PI because
it has been misclassified as a Chinstrap (orange).

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins_BlFl}

}

\caption[Checking what is learned from the cheem viewer]{Checking what is learned from the cheem viewer. This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI. A Gentoo (purple) misclassified as a Chinstrap (orange). The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
\end{figure}
\end{CodeChunk}

There is more separation visible in the attribution space than in the
data space, as would be expected. The predicted vs observed plot reveals
a handful of misclassified observations. A Gentoo that has been wrongly
labeled as a Chinstrap is selected for illustration. The PI is a
misclassified point (represented by the asterisk in the global view and
a dashed vertical line in the tour view). The CI is a correctly
classified point (represented by an \(\times\) and a vertical dotted
line).

The radial tour starts from the attribution projection of the
misclassified observation (b, left). The important variables identified
by SHAP in the (wrong) prediction for this observation are mostly
\texttt{bl} and \texttt{bd} with small contributions of \texttt{fl} and
\texttt{bm}. This projection is a view where the Gentoo (purple) looks
much more likely for this observation than Chinstrap. That is, this
combination of variables is not particularly useful because the PI looks
very much like other Gentoo penguins. The radial tour is used to vary
the contribution of flipper length (\texttt{fl}) to explore this. (In
our exploration, this was the third variable explored. It is typically
helpful to explore the variables with more significant contributions,
here \texttt{bl} and \texttt{bd}. Still, when doing this, nothing was
revealed about how the PI differed from other Gentoos). On varying
\texttt{fl} as it contributes increasingly to the projection (b, right),
more and more, this penguin looks like a Chinstrap. This suggests that
\texttt{fl} should be considered an important variable for explaining
the (wrong) prediction.

Figure \ref{fig:casepenguinsblfl} confirms that flipper length
(\texttt{fl}) is vital for the confusion of the PI as a Chinstrap. Here,
flipper length and body length are plotted, and the PI can be seen to be
closer to the Chinstrap group in these two variables, mainly because it
has an unusually low value of flipper length relative to other Gentoos.
From this view, it makes sense that it is a hard observation to account
for, as decision trees can only partition only vertical and horizontal
lines.

\hypertarget{chocolates-milkdark-classification}{%
\subsection{Chocolates, Milk/Dark
Classification}\label{chocolates-milkdark-classification}}

The chocolates data set consists of 88 observations of ten nutritional
measurements determined from their labels and labeled as either milk or
dark. Dark chocolate is considered healthier than milk. Students
collected the data during the Iowa State University class STAT503 from
nutritional information on the manufacturer's websites and were
normalized to 100g equivalents. The data is available in the
\textbf{cheem} package. A random forest model is used for the
classification of chocolate types.

It could be interesting to examine the nutritional properties of any
dark chocolates that have been misclassified as milk. A reason to do
this is that a dark chocolate, nutritionally more like milk should not
be considered a healthy alternative. It is interesting to explore which
nutritional variables contribute most to misclassification.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates}

}

\caption[Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green). From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation. So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber. When Sugar is zero, Fiber contributes strongly towards the left. In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber. This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate. (A video of the tour animation can be found at https://vimeo.com/666431143.)}\label{fig:casechocolates}
\end{figure}
\end{CodeChunk}

This type of exploration is shown in Figure \ref{fig:casechocolates},
where a chocolate labeled dark but predicted to be milk is chosen as the
PI (observation 22). It is compared with a CI that is a correctly
classified dark chocolate (observation 7). The PCA plot and the tree
SHAP PCA plots (a) show a big difference between the two chocolate types
but with confusion for a handful of observations. The misclassifications
are more apparent in the observed vs predicted plot and can be seen to
be mistaken in both ways: milk to dark and dark to milk.

The attribution projection for chocolate 22 suggests that Fiber, Sugars,
and Calories are most responsible for its incorrect prediction. The way
to read this plot is to see that Fiber has a large negative value while
Sugars and Calories have reasonably large positive values. In the
density plot, observations on the very left of the display would have
high values of Fiber (matching the negative projection coefficient) and
low values of Sugars and Calories. The opposite would be interpreting a
point with high values in this plot. The dark chocolates (orange) are
primarily on the left, and this is a reason why they are considered to
be healthier: high fiber and low sugar. The density of milk chocolates
is further to the right, indicating that they generally have low fiber
and high sugar.

The PI (dashed line) can be viewed against the CI (dotted line). Now,
one needs to pay attention to the parallel plot of the SHAP values,
which are local to a particular observation, and the density plot, which
is the same projection of all observations as specified by the SHAP
values of the PI. The variable contribution of the two different
predictions can be quickly compared in the parallel coordinate plot. The
PI differs from the comparison primarily on the Fiber variable, which
suggests that this is the reason for the incorrect prediction.

From the density plot, which is the attribution projection corresponding
to the PI, both observations are more like dark chocolates. Varying the
contribution of Sugars and altogether removing it from the projection is
where the difference becomes apparent. When a frame with contribution
primarily from Fiber is examined observation 22 looks more like a milk
chocolate.

It would also be interesting to explore an inverse misclassification. In
this case, a milk chocolate is selected while it was misclassified as a
dark chocolate. Chocolate 84 is just this case and is compared with a
correctly predicted milk chocolate (observation 71). The corresponding
global view and radial tour frames are shown in Figure
\ref{fig:casechocolatesinverse}.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/case_chocolates_inverse}

}

\caption[Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange). In the attribution projection, the PI could be either milk or dark. Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates. The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied. If Sodium had a larger contribution to the prediction (like in this view). the PI would look more like other milk chocolates. (A video of the tour animation can be found at https://vimeo.com/666431148.)}\label{fig:casechocolatesinverse}
\end{figure}
\end{CodeChunk}

The difference of position in the tree SHAP PCA with the previous case
is quite significant; this gives a higher-level sense that the
attributions should be quite different. Looking at the attribution
projection, this is found to be the case. Previously, Fiber was
essential while it is absent from the attribution in this case.
Conversely, Calories from Fat and Total Fat have high attributions here,
while they were unimportant in the preceding case.

Comparing the attribution with the CI (dotted line), large discrepancies
in Sodium and Fiber are identified. The contribution of Sodium is
selected to be varied. Even in the initial projection, the observation
looks slightly more like its observed milk than predicted dark
chocolate. The misclassification appears least supported when the basis
reaches sodium attribution of typical dark chocolate.

\hypertarget{fifa-wage-regression}{%
\subsection{FIFA, Wage Regression}\label{fifa-wage-regression}}

The 2020 season FIFA data \citep{leone_fifa_2020, biecek_dalex_2018}
contains many skill measurements of soccer/football players and wage
information. Nine higher-level skill groupings were identified and
aggregated from highly correlated variables. A random forest model is
fit from these predictors, regressing player wages {[}2020 euros{]}. The
model was fit from 5000 observations before being thinned to 500 players
to mitigate occlusion and render time. Continuing from the exploration
in Section \textbackslash ref\{sec:explanations), we are interested to
see the difference in attribution based on the exogenous player
position. That is, the model should be able to use multiple linear
profiles to better predict the wages from different field positions of
players despite not having this information. A leading offensive fielder
(L. Messi) is compared with a top defensive fielder (V. van Dijk). The
same observations were used in Figure \ref{fig:shapdistrbd}.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{./figures/case_fifa}

}

\caption[Exploring the wages relative to skill measurements in the FIFA 2020 data]{Exploring the wages relative to skill measurements in the FIFA 2020 data. Star offensive player (L. Messi) is the PI, and he is compared with a top defensive player (V. van Dijk). The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages. Defense (`def`) is the chosen variable to vary. It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot). The increased contribution in defense comes at the expense of offensive and reaction skills. The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the LVA. (A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
\end{figure}
\end{CodeChunk}

Figure \ref{fig:casefifa} tests the support of the LVA. Offensive and
reaction skills (\texttt{off} and \texttt{rct}) are both crucial to
explaining a star offensive player. If either of them were rotated out,
the other would be rotated into the frame, maintaining a far-right
position. However, increasing the contribution of a variable with low
importance would rotate both variables out of the frame.

The contribution from \texttt{def} will be varied to contrast with
offensive skills. As the contribution of defensive skills increases,
Messi's is no longer separated from the group. Players with high values
in defensive skills are now the rightmost points. In terms of what-if
analysis, the difference between the data mean and his predicted wages
would be halved if Messi's tree SHAP attributions were at these levels.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{./figures/case_ames2018}

}

\caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data. The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable. The PI has a higher predicted value than the CI in the attribution projection. Reducing the contribution of Lot Area brings these two prices in line. This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar. That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price. (A video showing the animation is at https://vimeo.com/666431134.)}\label{fig:caseames}
\end{figure}
\end{CodeChunk}

\hypertarget{ames-housing-2018-sales-price-regression}{%
\subsection{Ames Housing 2018, Sales Price
Regression}\label{ames-housing-2018-sales-price-regression}}

Ames housing data 2018 \citep{de_cock_ames_2011, prevek18_ames_2018} was
subset to North Ames (the neighborhood with the most house sales). The
remaining are 338 house sales. A random forest model was fit, predicting
the sale price {[}USD{]} from the property variables: Lot Area
(\texttt{LtA}), Overall Quality (\texttt{Qlt}), Year the house was Built
(\texttt{YrB}), Living Area (\texttt{LvA}), number of Bathrooms
(\texttt{Bth}), number of Bedrooms (\texttt{Bdr}), the total number of
Rooms (\texttt{Rms}), Year the Garage was Built (\texttt{GYB}), and
Garage Area (\texttt{GrA}). Using interactions with the global view, a
house with an extreme negative residual and an accurate observation with
a similar prediction is selected.

Figure \ref{fig:caseames} selects the house sale 74, a sizable under
prediction with an enormous Lot Area contribution. The CI has a similar
predicted price though the prediction was accurate and gives almost no
attribution to lot size. The attribution projection places observations
with high Living Areas to the right. The contribution of Living Area
contrasts the contribution of this variable. As the contribution of Lot
Area decreases, the predictive power decreases for the PI, while the CI
remains stationary. This large importance in the Living Area is
relatively uncommon. Boosting tree models may be more resilient to such
an under-prediction as they would up-weighting this residual and force
its inclusion in the final model.

\hypertarget{sec:cheemdiscussion}{%
\section{Discussion}\label{sec:cheemdiscussion}}

There is a clear need to extend the interpretability of black box
models. With techniques such as SHAP, LIME, Break-down, one can
calculate LEs, i.e.~for every observation in the data. These techniques
quantify for each observation how strongly particular variables affect
the model's predictions. Surprisingly few techniques allow us to
understand the global distribution of these LEs. Unsupervised data
exploration techniques applied to data show how useful they are for
identifying outliers, identifying clusters of observations or
discovering correlations between variables. All of these tasks can be
performed for a set of explanations.

To address this challenge in this paper provides a technique that builds
on LEs to explore the variable importance local to an observation. The
LVA is converted into an attribution projection from which variable
contributions are varied using a radial tour. Several diagnostic plots
are provided to assist with understanding the sensitivity of the
prediction to particular variables. A global view shows the data space,
explanation space, and residual plot. The user can interactively select
observations to compare, contrast, and study further. Then the radial
tour is used to explore the variable sensitivity identified by the
attribution projection.

This approach has been illustrated using four data examples of random
forest models with the tree SHAP LVA. LEs focus on the model fit, and
help to dissect which variables are most responsible for the fitted
value. They can also form the basis of learning how the model has got it
wrong, with when the observation is misclassified or has a large
residual.

In the penguins example, we showed how the misclassification of a
penguin arose due to it having an unusually small flipper size compared
to others of its species. This was verified by making a follow-up plot
of the data. The chocolates example shows how a dark chocolate was
misclassified primarily due to its attribution to Fiber, and a milk
chocolate was misclassified as dark due to its lowish Sodium value. In
the FIFA example, we show how low Messi's salary would be if it depended
on defensive skills. In the Ames housing data, an inaccurate prediction
for a house was likely due to the Lot Area is not being effectively used
by the random forest model.

This analysis is manually intensive and thus only feasible for
investigating a few observations. The recommended approach is to
investigate an observation where the model has not predicted accurately
and compare it with an observation with similar predictor values where
model fitted well. The radial tour launches from the attribution
projection to enable exploration of the sensitivity of the prediction to
any variable. It can be helpful to make additional plots of the
variables and responses to cross-check interpretations made from the
cheem viewer. This methodology provides an additional tool in the box
for studying model fitting.

\hypertarget{sec:infrastructure}{%
\section{Package Infrastructure}\label{sec:infrastructure}}

An implementation is provided in the open-source \textbf{R} package
\textbf{cheem}, available on CRAN at
\url{https://CRAN.R-project.org/package=cheem}. Example data sets are
provided, and you can upload your data after model fitting and computing
the LVAs. The LVAs need to be pre-computed and uploaded. Examples show
how to do this for tree SHAP values, using \textbf{treeshap} (tree based
models from \textbf{gbm}, \textbf{lightgbm}, \textbf{randomForest},
\textbf{ranger}, or \textbf{xgboost} \citet{greenwell_gbm_2020};
\citet{shi_lightgbm_2022}; \citet{liaw_classification_2002};
\citet{wright_ranger_2017}; \citet{chen_xgboost_2021}, respectively).
The SHAP and oscillation explanations could be easily added using
\texttt{DALEX::explain()}
\citep{biecek_dalex_2018, biecek_explanatory_2021}.

The application was made with \textbf{shiny} \citep{chang_shiny_2021}.
The tour visual is built with \textbf{spinifex}
\citep{spyrison_spinifex_2020}. Both views are created first with
\textbf{ggplot2} \citep{wickham_ggplot2_2016} and then rendered as
interactive \texttt{html} widgets with \textbf{plotly}
\citep{sievert_interactive_2020}. \textbf{DALEX}
\citep{biecek_dalex_2018} and \emph{Explanatory Model Analysis}
\citep{biecek_explanatory_2021} are helpfule for understanding LEs and
how to apply them.

The package can be installed from CRAN, and the application can be run
using the following \textbf{R} code:

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("cheem", dependencies = TRUE)
R> library("cheem")
R> run_app()
\end{CodeInput}
\end{CodeChunk}

Alternatively,

\begin{itemize}
\tightlist
\item
  A version of the cheem viewer shiny app can be directly accessed at
  \url{https://ebsmonash.shinyapps.io/cheem_initial/}.
\item
  The development version of the package is available at
  \url{https://github.com/nspyrison/cheem}, and
\item
  Documentation of the package can be found at
  \url{https://nspyrison.github.io/cheem/}.
\end{itemize}

Follow the examples provided with the package to compute the LVAs (using
\texttt{?cheem\_ls}). The application expects the output returned by
\texttt{cheem\_ls()}, saved to an \texttt{rds} file with
\texttt{saveRDS()} to be uploaded.

\hypertarget{acknowledgments}{%
\subsection*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{subsection}{Acknowledgments}

Kim Marriott provided advice on many aspects of this work, especially on
the explanations in the applications section. This research was
supported by the Australian Government Research Training Program (RTP)
scholarships. Thanks to Jieyang Chong for helping proofread this
article. The namesake, Cheem, refers to a fictional race of humanoid
trees from Doctor Who lore. \textbf{DALEX} pulls on from that universe,
and we initially apply tree SHAP explanations specific to tree-based
models.

\renewcommand\refname{References}
\bibliography{paper.bib}

\end{document}